---
title: "Using MARSS"
output: html_notebook
---

In this document, I use MARSS package to analyze the same data (Indonesia nest counts of leatherback turtles) to see how the analysis will reveal.  Shouldn't be much different but curious about how missing data are imputed. 

Get libraries and bring in my functions for this project. 

```{r}
rm(list=ls())
library(MARSS)
library(tidyverse)
source("Dc_Indonesia_nesting_fcns.R")

run.date <- Sys.Date()

year.begin <- 1999
year.end <- 2018

fill.color <-  "darkseagreen"
fill.color.summer <- "darksalmon"
fill.color.winter <- "gray65"
fill.alpha <-  0.65
line.color <-  "darkblue"
line.color.summer <- "red3"
line.color.winter <- "greenyellow"
data.color <- "black"
data.size <- 1.5
obsd.color <- "red2"

obsd.size <- 1

locations <- c("JM", "W")
```


Extract the data for Jamursba-Medi and Wermon and combine them together for use in MARSS.  MARSS inputs require time in columns (n x T matrix):
```{r}
data.0.JM <- data.extract(location = "JM", year.begin = 1981, year.end = 2019)

data.0.JM$data.1 %>% 
  mutate(f_month = as.factor(Month),
         f_year = as.factor(Year),
         Season = ifelse(Month < 4, Year - 1, Year),
         season = ifelse(Month > 3 & Month < 10, 
                         "summer", "winter")) -> data.1.JM

data.1.JM.1999 <- filter(data.1.JM, Year > 1998)


data.0.W <- data.extract(location = "W", year.begin = 2002, year.end = 2019)

data.0.W$data.1 %>% 
  mutate(f_month = as.factor(Month),
         f_year = as.factor(Year),
         Season = ifelse(Month < 4, Year - 1, Year),
         season = ifelse(Month > 3 & Month < 10, 
                         "summer", "winter")) -> data.1.W

data.1.W %>% filter(Year > 2002) -> data.1.W.2003

data.1.JM.1999 %>% left_join(data.1.W.2003, by = c("Year", "Month")) %>%
  select(Year, Month, Frac.Year.x, Season.x, season.x, Nests.x, Nests.y) %>%
  transmute(Year = Year, Month = Month, Frac.Year = Frac.Year.x,
         Season = Season.x, season = season.x, 
         Nests.JM = Nests.x, Nests.W = Nests.y)-> data.1.JM.W

data.MARSS <- log(t(select(data.1.JM.W, Nests.JM, Nests.W)))
```

I follow the user guide, starting around p. 35. 

Try fitting a simple model. The default model in MARSS is the following.

x[t] = I x[t-1] + u + w[t], w[t] ~ MVN(0, Q)
y[t] = I x[t] + 0 + v[t], v[t] ~ MVN(0, R)

Q and R are time-independent.  0 is a zero vector, u is time-independent. 

maxit had to be increased to reach Abstol convergence. log-log convergence did not reach at maxit = 10000. Seems to be the main problem is R.diag - the observation variance parameter. Even though, there is just one... 
```{r}
maxit <- 10000
if (!file.exists(paste0("RData/MARSS_fit1_", run.date, ".rds"))){
  fit.1 <- MARSS(data.MARSS, 
                 control = list(maxit = maxit),
                 method = "BFGS")
  
  saveRDS(fit.1, file = paste0("RData/MARSS_fit1_", run.date, ".rds"))
  
} else {
  fit.1 <- readRDS(file = paste0("RData/MARSS_fit1_", run.date, ".rds"))
}

```

The "BFGS" method works faster and convergence reaches. When "method" is left as its default ("kem"), a lot slower and it does not converge... 

The model does not converge - not a huge surprise. 

x[t] = B x[t-1] + w[t], w[t] ~ MVN(0, Q)
y[t] = I x[t] + v[t], v[t] ~ MVN(0, R)

Change Q to unconstratined so that it's not just diagonal qs.
Also add independent slope to each beach.
```{r}
if (!file.exists(paste0("RData/MARSS_fit2_", run.date, ".rds"))){
  fit.2 <- MARSS(data.MARSS, 
                 model = list(Q = "unconstrained",
                              B = "diagonal and unequal",
                              R = "diagonal and equal"),
                 control = list(maxit = maxit),
                 method = "BFGS")

  saveRDS(fit.2, file = paste0("RData/MARSS_fit2_", run.date, ".rds"))
  
} else {
  fit.2 <- readRDS(file = paste0("RData/MARSS_fit2_", run.date, ".rds"))
}

```

Convergence issue persists... not very surprising given our seasonal changes in counts. 

Change Q to equal variance and covariance. This does not run with method = "BFGS."  The error message states that when using BFGS, var-cov matrices must be either diagonal or uncontrained at t=1.

```{r}
if (!file.exists(paste0("RData/MARSS_fit3_", run.date, ".rds"))){
  fit.3 <- MARSS(data.MARSS, 
                 model = list(Q = "equalvarcov"),
                 control = list(maxit = 5000),
                 method = "kem")
  
  
  saveRDS(fit.3, file = paste0("RData/MARSS_fit3_", run.date, ".rds"))
  
} else {
  fit.3 <- readRDS(file = paste0("RData/MARSS_fit3_", run.date, ".rds"))
}
```

Alternatively, we can use "season" as a state covariate.  


x[t] = B x[t-1] + C[t] c[t] + w[t], w[t] ~ MVN(0, Q)
y[t] = Z x[t] + v[t], v[t] ~ MVN(0, R)

Month is not working so switch to that... I get the followinbg message when using "month":

Stopped at iter=2 in MARSSkem at U update. denom is not invertible.
 This means some of the U (+ C) terms cannot be estimated.
 Type MARSSinfo('denominv') for more info. 
 par, kf, states, iter, loglike are the last values before the error.
 

```{r}

c.in <- matrix(0, nrow = 2, ncol = nrow(data.1.JM.W))
c.in[1, data.1.JM.W$season == "summer"] <- 1
c.in[2, data.1.JM.W$season == "winter"] <- 1

if (!file.exists(paste0("RData/MARSS_fit4_", run.date, ".rds"))){
  fit.4 <- MARSS(data.MARSS,
                 model = list(B = "diagonal and unequal",
                              Q = "diagonal and unequal",
                              Z = "identity", 
                              R = "diagonal and equal",
                              D = "zero",
                              d = "zero",
                              C = "diagonal and unequal",
                              c = c.in),
                 control = list(maxit = 5000,
                                trace = 1),
                 method = "BFGS")
  
  saveRDS(fit.4, file = paste0("RData/MARSS_fit4_", run.date, ".rds"))
  
} else {
  fit.4 <- readRDS(file = paste0("RData/MARSS_fit4_", run.date, ".rds"))
}



```

How can I add seasonal and location specific state covariates defined?
Change C to "unconstrained"?  Then the "not invertible" error returns... We could have one parameter between two locations for summer and different for winter? It goes a few more steps before turning into "not invertible" error again.  

Changing the optimization method to BFGS made it to work. 
```{r}
# keep the state covariate
c.in <- matrix(0, nrow = 2, ncol = nrow(data.1.JM.W))
c.in[1, data.1.JM.W$season == "summer"] <- 1
c.in[2, data.1.JM.W$season == "winter"] <- 1
rownames(c.in ) <- c("summer", "winter")

# but make C to be unconstrained to have season and location specific parameters. 

if (!file.exists(paste0("RData/MARSS_fit5_", run.date, ".rds"))){
  fit.5 <- MARSS(data.MARSS,
                 model = list(B = "diagonal and unequal",
                              Q = "diagonal and unequal",
                              Z = "identity", 
                              R = "diagonal and equal",
                              D = "zero",
                              d = "zero",
                              C = "unconstrained",
                              c = c.in),
                 control = list(maxit = 5000,
                                trace = 1),
                 method = "BFGS")
  
  saveRDS(fit.5, file = paste0("RData/MARSS_fit5_", run.date, ".rds"))
  
} else {
  fit.5 <- readRDS(file = paste0("RData/MARSS_fit5_", run.date, ".rds"))
}

```

What if we add month as a covariate for observations? Add month as a covariate. See p. 167 (13.6.1). 

x[t] = B x[t-1] + C[t] c[t] + w[t], w[t] ~ MVN(0, Q)
y[t] = Z x[t] + D[t] d[t] + v[t], v[t] ~ MVN(0, R)

```{r}

# state covariate - month
period <- 12
d.in <- diag(period)

TT <- dim(data.MARSS)[2]
for (i in 2:ceiling(TT/period)){
  d.in <- cbind(d.in, diag(period))
}

rownames(d.in ) <- month.abb

# 
D.mat <- matrix(month.abb, 2, 12, byrow = TRUE)

c.in <- matrix(0, nrow = 2, ncol = nrow(data.1.JM.W))
c.in[1, data.1.JM.W$season == "summer"] <- 1
c.in[2, data.1.JM.W$season == "winter"] <- 1
rownames(c.in ) <- c("summer", "winter")

if (!file.exists(paste0("RData/MARSS_fit6_", run.date, ".rds"))){
  fit.6 <- MARSS(data.MARSS,
                 model = list(B = "diagonal and unequal",
                              Q = "diagonal and unequal",
                              Z = "identity", 
                              R = "diagonal and equal",
                              D = "unconstrained",
                              d = d.in,
                              C = "unconstrained",
                              c = c.in),
                 control = list(maxit = maxit,
                                trace = 1),
                 method = "BFGS")
  
  saveRDS(fit.6, file = paste0("RData/MARSS_fit6_", run.date, ".rds"))
  
} else {
  fit.6 <- readRDS(file = paste0("RData/MARSS_fit6_", run.date, ".rds"))
}


```

With a smaller tolerance ( conv.test.slope.tol = 0.05), a few parameters did not converge.  It was fine when the tolerance was set at 0.1.  Changing the optimization method to BFGS seemed to fix the problem.

Try Fourier series for the seasonal counts - I'll try the state space only first. (p. 170)

```{r}
period <- 12

cos.t <- cos(2 * pi * seq(TT) /  period)
sin.t <- sin(2 * pi * seq(TT) /  period)

c.Four <- rbind(cos.t, sin.t)

if (!file.exists(paste0("RData/MARSS_fit7_", run.date, ".rds"))){
  fit.7 <- MARSS(data.MARSS,
                 model = list(B = "diagonal and unequal",
                              Q = "diagonal and unequal",
                              Z = "identity", 
                              R = "diagonal and equal",
                              D = "zero",
                              d = "zero",
                              C = "unconstrained",
                              c = c.Four),
                 control = list(maxit = maxit,
                                trace = 1),
                 method = "BFGS")
  
  saveRDS(fit.7, file = paste0("RData/MARSS_fit7_", run.date, ".rds"))
  
} else {
  fit.7 <- readRDS(file = paste0("RData/MARSS_fit7_", run.date, ".rds"))
}
```

Again, R.diag did not converge... so why not add seasonal effects on observation. Changing the optimization method fixed the problem. 

The next model uses a discrete Fourier series to model the process and seasonal effects on observation. Note that I set 12 months as a period - this may not be good as we have two nesting seasons in Wermon.  

```{r}
d.in <- matrix(0, nrow = 2, ncol = nrow(data.1.JM.W))
d.in[1, data.1.JM.W$season == "summer"] <- 1
d.in[2, data.1.JM.W$season == "winter"] <- 1

rownames(d.in) <- c("summer", "winter")

period <- 12

cos.t <- cos(2 * pi * seq(TT) /  period)
sin.t <- sin(2 * pi * seq(TT) /  period)

c.Four <- rbind(cos.t, sin.t)

if (!file.exists(paste0("RData/MARSS_fit8_", run.date, ".rds"))){
  fit.8 <- MARSS(data.MARSS,
                 model = list(B = "diagonal and unequal",
                              Q = "diagonal and unequal",
                              Z = "identity", 
                              R = "diagonal and equal",
                              D = "unconstrained",
                              d = d.in,
                              C = "unconstrained",
                              c = c.Four),
                 control = list(maxit = maxit,
                                trace = 1),
                 method = "BFGS")
  
  saveRDS(fit.8, file = paste0("RData/MARSS_fit8_", run.date, ".rds"))
  
} else {
  fit.8 <- readRDS(file = paste0("RData/MARSS_fit8_", run.date, ".rds"))
}

```

R.diag does not converge with this model... BFGS fixed the problem. 

Because Wermon has two peaks in nesting, I try to add that into the discrete Fourier series for Wermon. I can't figure out how to have two different periods for the two locations... For this one, I just use 6 months as the common period - not quite right... 2019-04-09

```{r}
d.in <- matrix(0, nrow = 2, ncol = nrow(data.1.JM.W))
d.in[1, data.1.JM.W$season == "summer"] <- 1
d.in[2, data.1.JM.W$season == "winter"] <- 1

rownames(d.in) <- c("summer", "winter")

period <- 6

cos.t <- cos(2 * pi * seq(TT) /  period)
sin.t <- sin(2 * pi * seq(TT) /  period)

c.Four <- rbind(cos.t, sin.t)

if (!file.exists(paste0("RData/MARSS_fit9_", run.date, ".rds"))){
  fit.9 <- MARSS(data.MARSS,
                 model = list(B = "diagonal and unequal",
                              Q = "diagonal and unequal",
                              Z = "identity", 
                              R = "diagonal and equal",
                              D = "unconstrained",
                              d = d.in,
                              C = "unconstrained",
                              c = c.Four),
                 control = list(maxit = maxit,
                                trace = 1),
                 method = "BFGS")
  
  saveRDS(fit.9, file = paste0("RData/MARSS_fit9_", run.date, ".rds"))
  
} else {
  fit.9 <- readRDS(file = paste0("RData/MARSS_fit9_", run.date, ".rds"))
}

```


So, now we have results from eight models.  Compare them using AICc/AIC.

```{r}
data.frame(Model = c("M1", "M2", "M3", "M4",
                     "M5", "M6", "M7", "M8", 
                     "M9"),
           AICc = round(c(fit.1$AICc, fit.2$AICc, fit.3$AICc, fit.4$AICc,
                          fit.5$AICc, fit.6$AICc, fit.7$AICc, fit.8$AICc,
                          fit.9$AICc), 2)) %>%
  arrange(AICc)
```

Model 9 is by far the best according to AICc.  Take a look at predicted values.

```{r}
C.M9 <- coef(fit.9, type = "matrix")$C
seasonal.effect <- C.M9 %*% fit.9$call$model$c[, 1:(2*period)]

rownames(seasonal.effect) <- locations
colnames(seasonal.effect) <- 1:(period*2)

seasonal.effect.df <- data.frame(t(seasonal.effect)) %>% 
  rownames_to_column(var = "month") %>%
  reshape::melt(id = "month", 
                variable_name = "location") %>%
  mutate(month = as.numeric(month))
  
 
ggplot(data = seasonal.effect.df) + 
  geom_line(aes(x = month, 
                y = value, 
                color = location))
```

Then look at residuals.  Start here... 


