---
title: "Nest count imputation using a discrete Fourier series - Wermon"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
source("Dc_Indonesia_nesting_fcns.R")
run.date <- "2019-04-30"
save.fig <- F

models.df <- model.names()
year.begin <- 2003
year.end <- 2019

fill.color <-  "darkseagreen"
fill.color.summer <- "darksalmon"
fill.color.winter <- "gray65"
fill.alpha <-  0.65
line.color <-  "darkblue"
line.color.summer <- "red3"
line.color.winter <-"red3"  #"greenyellow"
data.color <- "black"
data.size <- 1.5
obsd.color <- "red2"

obsd.size <- 1
```

This document describes how leatherback turtle nest count data were imputed for Jamursba-Medi. All models were ran separately and results stored in .rds files. 

We use the Bayesian state-space approach, where the mean of the state process is modeled with an auto-regressive model with time lag of one step (AR(1)). 

M[t] = S[t] * X[t-1]

where s[t] is the “slope” parameter between two consecutive counts (transformed into the natural log scale). When s[t] > 1, the counts increase whereas s[t] < 1, they decline. This slope parameter is modeled with a discrete Fourier series with 6 months as a period. 

s[k] = s1 \* cos(2 \*  pi \* k/6) + s2 \* sin(2 \* pi \* k/6), where k = 1, ..., 12 corresponding to months within each nesting season (starts April and ends March).

The state-space is modeled with a normal distribution.

X[t] ~ N(M[t], v[t]),

where v[t] is the variance of X[t]. The variance also can be modeled in a few different scenarios as was in the slope parameter. It can be constant (v[t] = v for all t), month-specific (v[month]), or constant CV (sqrt(V[t])/M[t] = constant). 

Given the state process, observations are modeled with normal or t distributions. 

Y[t] ~ N(X[t], v.obs)

Y[t] ~ t(X[t], v.obs, df)

where v.obs is the observation variance and df is the degrees-of-freedom parameter of the t distribution. The observation variance is assumed to be constant throughout the dataset. In total 8 models were developed and predictive accuracy was compared to select the most accurate model (Table x).  Model selection process is explained in the subsequent section.  The analysis is conducted in the natural logarithm scale. We define the following:

x[t] = ln(X[t])

m[t] = ln(M[t])

y[t] = ln(Y[t])

s[t] = ln(S[t])

The models were fit to datasets from turtle nesting beaches. Bayesian computations were conducted using jags through jagsUI package in R.  

##Model 1
First model is the simplest; the state space is modeled with Normal distribution and observation is modeled also with normal distribution. Variance is assumed constant:

m[t] = s + x[t-1]

x[t] ~ N(m[t], v.pro)

y[t] ~ N(x[t], v.obs)

```{r model1, cache=TRUE, include=TRUE}
model.name <- "model_SSAR1_logY_norm_norm_theta_Four.txt"
str.root <- strsplit(strsplit(model.name, "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_W_", 
                    year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M1.loo.out <- M$loo.out$loo.out
M1.jm <- M$jags.out
M1.Xs.stats <- M$Xs.stats
M1.ys.stats <- M$ys.stats
```

Diagnostics of the expected log predictive density indicated that the model was not appropriate, where many (```r signif(sum(M1.loo.out$diagnostics$pareto_k > 0.7)/length(M1.loo.out$diagnostics$pareto_k) * 100, 3)``` %) of estimated Pareto k statistics (a measure of model fit - TRUE?) were greater than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M1.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M1.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M1.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M1, fig.cap = "Figure. PSIS diagnostic plot for M1"}

plot(M1.loo.out)
```

```{r}
pareto.k <- M1.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.data$y)

pareto.df.M1 <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df.M1) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))


```


```{r}
p.pareto.M1 <- ggplot(data = pareto.df.M1) +  
  geom_point(aes(x = datapoint, 
                 y = khat,
                 color = -khat)) + 
  geom_hline(yintercept = 0.5,
             color = "red",
             size = 0.5) +
  geom_hline(yintercept = 0.7,
             color = "red",
             size = 0.7) +
  geom_hline(yintercept = 1.0,
             color = "red",
             size = 1.0) +
  theme(legend.position = "none") + 
  labs(x = "Data point", y = "Pareto shape k",
       title = "PSIS diagnostic plot (Model 1)")

if (save.fig)
  ggsave(p.pareto.M1, filename = "figures/W_PSIS_M1.png",
         device = "png", dpi = 600)

print(p.pareto.M1)
```



```{r}

M$ys.stats %>% rowid_to_column() %>% 
  mutate(pareto_k = NA) -> ys.stats

noNA.row <- ys.stats$rowid[!is.na(ys.stats$obsY)]

ys.stats[noNA.row, "pareto_k"] <- pareto.k

ggplot(data = ys.stats) +   
  geom_path(aes(x = time, y = obsY), 
            alpha = 0.5) +
  geom_point(aes(x = time, y = obsY, 
                 color = pareto_k),
             size = 3) +
  scale_color_viridis_c(limits = c(0.0, 1.3))

  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))



```

##Model 2
In the second model, the state variance parameter was modeled as a function of the mean and constant CV.

m[t] = s + x[t-1] 

x[t] ~ N(m[t], v.pro[t]), sqrt(v.pro[t]) = m[t] * CV

y[t] ~ N(x[t], v.obs)

```{r model2, cache=TRUE, include=TRUE}
model.name <- "model_SSAR1_logY_norm_norm_theta_Four_constCV.txt"
str.root <- strsplit(strsplit(model.name, "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_W_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M2.loo.out <- M$loo.out$loo.out
M2.jm <- M$jags.out
M2.Xs.stats <- M$Xs.stats
M2.ys.stats <- M$ys.stats

```

Although better than the first model, some estimated Pareto k statistics were greater than 0.7 (```r signif(sum(M2.loo.out$diagnostics$pareto_k > 0.7)/length(M1.loo.out$diagnostics$pareto_k) * 100, 3)``` %). Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M2.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M2.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M2.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M2, fig.cap="Figure. PSIS diagnostic plot for M2"}
plot(M2.loo.out)
```


```{r}
pareto.k <- M2.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.data$y)

pareto.df.M2 <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df.M2) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

```{r}
p.pareto.M2 <- ggplot(data = pareto.df.M2) +  
  geom_point(aes(x = datapoint, 
                 y = khat,
                 color = -khat)) + 
  geom_hline(yintercept = 0.5,
             color = "red",
             size = 0.5) +
  geom_hline(yintercept = 0.7,
             color = "red",
             size = 0.7) +
  geom_hline(yintercept = 1.0,
             color = "red",
             size = 1.0) +
  theme(legend.position = "none") + 
  labs(x = "Data point", y = "Pareto shape k",
       title = "PSIS diagnostic plot (Model 2)")

if (save.fig)
  ggsave(p.pareto.M2, filename = "figures/W_PSIS_M2.png",
         device = "png", dpi = 600)

print(p.pareto.M2)
```

```{r}

M$ys.stats %>% rowid_to_column() %>% 
  mutate(pareto_k = NA) -> ys.stats

noNA.row <- ys.stats$rowid[!is.na(ys.stats$obsY)]

ys.stats[noNA.row, "pareto_k"] <- pareto.k

ggplot(data = ys.stats) +   
  geom_path(aes(x = time, y = obsY), 
            alpha = 0.5) +
  geom_point(aes(x = time, y = obsY, 
                 color = pareto_k),
             size = 3) +
  scale_color_viridis_c(limits = c(0.0, 1.3))

  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))



```


##Model 3
The third model has  month-specific variances.

m[t] = s + x[t-1]

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v.pro[month], month = 1, 2, ..., 12

y[t] ~ N(x[t], v.obs)

```{r model3, cache=TRUE, include=TRUE}
model.name <- "model_SSAR1_logY_norm_norm_theta_Four_varM.txt"
str.root <- strsplit(strsplit(model.name, "model_")[[1]][2], ".txt")[[1]]
M3 <- readRDS(paste0("RData/jagsout_", str.root, "_W_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M3.loo.out <- M$loo.out$loo.out
M3.jm <- M$jags.out
M3.Xs.stats <- M$Xs.stats
M3.ys.stats <- M$ys.stats

```

Some estimated Pareto k statistics were greater than 0.7 (```r signif(sum(M1.loo.out$diagnostics$pareto_k > 0.7)/length(M3.loo.out$diagnostics$pareto_k) * 100, 3)``` %). Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M3.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M3.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M3.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M3, fig.cap="Figure. PSIS diagnostic plot for M3"}
plot(M3.loo.out)
```

```{r}
pareto.k <- M3.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.data$y)

pareto.df.M3 <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df.M3) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```


```{r}
p.pareto.M3 <- ggplot(data = pareto.df.M3) +  
  geom_point(aes(x = datapoint, 
                 y = khat,
                 color = -khat)) + 
  geom_hline(yintercept = 0.5,
             color = "red",
             size = 0.5) +
  geom_hline(yintercept = 0.7,
             color = "red",
             size = 0.7) +
  geom_hline(yintercept = 1.0,
             color = "red",
             size = 1.0) +
  theme(legend.position = "none") + 
  labs(x = "Data point", y = "Pareto shape k",
       title = "PSIS diagnostic plot (Model 3)")

if (save.fig)
  ggsave(p.pareto.M3, filename = "figures/W_PSIS_M3.png",
         device = "png", dpi = 600)

print(p.pareto.M3)
```

```{r}

M$ys.stats %>% rowid_to_column() %>% 
  mutate(pareto_k = NA) -> ys.stats

noNA.row <- ys.stats$rowid[!is.na(ys.stats$obsY)]

ys.stats[noNA.row, "pareto_k"] <- pareto.k

ggplot(data = ys.stats) +   
  geom_path(aes(x = time, y = obsY), 
            alpha = 0.5) +
  geom_point(aes(x = time, y = obsY, 
                 color = pareto_k),
             size = 3) +
  scale_color_viridis_c(limits = c(0.0, 1.3))

  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))



```

##Model 4

Model 4 is similar to the first model but uses a different distribution for the state model; the observation is  modeled with t distribution. No seasonal change in variance or slope is considered. 

m[t] = s + x[t-1]

x[t] ~ N(m[t], v.pro)

y[t] ~ t(x[t], v.obs, df)

```{r model10, cache=TRUE, include=TRUE}
model.name <- "model_SSAR1_logY_norm_t_theta_Four.txt"
str.root <- strsplit(strsplit(model.name, "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_W_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M4.loo.out <- M$loo.out$loo.out
M4.jm <- M$jags.out
M4.Xs.stats <- M$Xs.stats
M4.ys.stats <- M$ys.stats

```

There were some large estimated Pareto k parameter values, where ```r sum(M4.loo.out$diagnostics$pareto_k > 0.7)/length(M4.loo.out$diagnostics$pareto_k) * 100```% were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M4.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M4.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M4.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M4, fig.cap="Figure. PSIS diagnostic plot for M4"}
plot(M4.loo.out)
```

```{r}
pareto.k <- M4.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.data$y)

pareto.df.M4 <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df.M4) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red")) 

```


```{r}
p.pareto.M4 <- ggplot(data = pareto.df.M4) +  
  geom_point(aes(x = datapoint, 
                 y = khat,
                 color = -khat)) + 
  geom_hline(yintercept = 0.5,
             color = "red",
             size = 0.5) +
  geom_hline(yintercept = 0.7,
             color = "red",
             size = 0.7) +
  geom_hline(yintercept = 1.0,
             color = "red",
             size = 1.0) +
  theme(legend.position = "none") + 
  labs(x = "Data point", y = "Pareto shape k",
       title = "PSIS diagnostic plot (Model 4)")

if (save.fig)
  ggsave(p.pareto.M4, filename = "figures/W_PSIS_M4.png",
         device = "png", dpi = 600)

print(p.pareto.M4)
```

```{r}

M$ys.stats %>% rowid_to_column() %>% 
  mutate(pareto_k = NA) -> ys.stats

noNA.row <- ys.stats$rowid[!is.na(ys.stats$obsY)]

ys.stats[noNA.row, "pareto_k"] <- pareto.k

ggplot(data = ys.stats) +   
  geom_path(aes(x = time, y = obsY), 
            alpha = 0.5) +
  geom_point(aes(x = time, y = obsY, 
                 color = pareto_k),
             size = 3) +
  scale_color_viridis_c(limits = c(0.0, 1.3))

  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))



```

##Model 5

The next model contained constant CV and t distribution for observations. 

m[t] = s + x[t-1] 

x[t] ~ N(m[t], v.pro[t]), sqrt(v.pro[t]) = m[t] * CV

y[t] ~ t(x[t], v.obs, df)

```{r model5, cache=TRUE, include=TRUE}
model.name <- "model_SSAR1_logY_norm_t_theta_Four_constCV.txt"
str.root <- strsplit(strsplit(model.name, "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_W_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M5.loo.out <- M$loo.out$loo.out
M5.jm <- M$jags.out
M5.Xs.stats <- M$Xs.stats
M5.ys.stats <- M$ys.stats

```

Model 5 showed better results with repsect to estimated Pareto k statistics, where (```r signif(sum(M5.loo.out$diagnostics$pareto_k > 0.7)/length(M5.loo.out$diagnostics$pareto_k) * 100, 3)``` % were greater than 0.7). Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M5.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M5.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M5.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M5, fig.cap="Figure. PSIS diagnostic plot for M5"} 
plot(M5.loo.out)
```

```{r}
pareto.k <- M5.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.data$y)

pareto.df.M5 <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df.M5) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```


```{r}
p.pareto.M5 <- ggplot(data = pareto.df.M5) +  
  geom_point(aes(x = datapoint, 
                 y = khat,
                 color = -khat)) + 
  geom_hline(yintercept = 0.5,
             color = "red",
             size = 0.5) +
  geom_hline(yintercept = 0.7,
             color = "red",
             size = 0.7) +
  geom_hline(yintercept = 1.0,
             color = "red",
             size = 1.0) +
  theme(legend.position = "none") + 
  labs(x = "Data point", y = "Pareto shape k",
       title = "PSIS diagnostic plot (Model 5)")

if (save.fig)
  ggsave(p.pareto.M5, filename = "figures/W_PSIS_M5.png",
         device = "png", dpi = 600)

print(p.pareto.M5)
```
```{r}

M$ys.stats %>% rowid_to_column() %>% 
  mutate(pareto_k = NA) -> ys.stats

noNA.row <- ys.stats$rowid[!is.na(ys.stats$obsY)]

ys.stats[noNA.row, "pareto_k"] <- pareto.k

ggplot(data = ys.stats) +   
  geom_path(aes(x = time, y = obsY), 
            alpha = 0.5) +
  geom_point(aes(x = time, y = obsY, 
                 color = pareto_k),
             size = 3) +
  scale_color_viridis_c(limits = c(0.0, 1.3))

  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))



```

##Model 6

Model 12 has a constant slope and month-specific variances. The model is as follows:

m[t] = s + x[t-1] 

x[t] ~ N(m[t], v.pro[t]), where v.pro[t] = v.pro[month], month = 1, 2, ..., 12

y[t] ~ t(x[t], v.obs, df)


```{r model6, cache=T, include=TRUE}
model.name <- "model_SSAR1_logY_norm_t_theta_Four_varM.txt"
str.root <- strsplit(strsplit(model.name, "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_W_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M6.loo.out <- M$loo.out$loo.out
M6.jm <- M$jags.out
M6.Xs.stats <- M$Xs.stats
M6.ys.stats <- M$ys.stats

```

The vast majority of estimated Pareto k parameter was greater than 0.7 (```r sum(M6.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M6.loo.out$diagnostics$pareto_k)```. Convergence was reached for all monitored parameters (minimum = ```r signif(min(unlist(lapply(M6.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M6.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M6.jm$Rhat, FUN = max, na.rm = T))), 3)```).  

```{r PSIS_M6, fig.cap="Figure. PSIS diagnostic plot for M6"}
plot(M6.loo.out)
```

```{r}
pareto.k <- M6.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.data$y)

pareto.df.M6 <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df.M6) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red")) 

```


```{r}
p.pareto.M6 <- ggplot(data = pareto.df.M6) +  
  geom_point(aes(x = datapoint, 
                 y = khat,
                 color = -khat)) + 
  geom_hline(yintercept = 0.5,
             color = "red",
             size = 0.5) +
  geom_hline(yintercept = 0.7,
             color = "red",
             size = 0.7) +
  geom_hline(yintercept = 1.0,
             color = "red",
             size = 1.0) +
  theme(legend.position = "none") + 
  labs(x = "Data point", y = "Pareto shape k",
       title = "PSIS diagnostic plot (Model 6)")

if (save.fig)
  ggsave(p.pareto.M6, filename = "figures/W_PSIS_M6.png",
         device = "png", dpi = 600)

print(p.pareto.M6)
```

```{r}

M$ys.stats %>% rowid_to_column() %>% 
  mutate(pareto_k = NA) -> ys.stats

noNA.row <- ys.stats$rowid[!is.na(ys.stats$obsY)]

ys.stats[noNA.row, "pareto_k"] <- pareto.k

ggplot(data = ys.stats) +   
  geom_path(aes(x = time, y = obsY), 
            alpha = 0.5) +
  geom_point(aes(x = time, y = obsY, 
                 color = pareto_k),
             size = 3) +
  scale_color_viridis_c(limits = c(0.0, 1.3))

  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))



```

#Model Comparison
First, we compared all models using using ELPD values although Pareto k statistics indicated that models 5, 4, 2, and 3 had acceptable predictive performance. 

```{r ModelComparison, include=TRUE}

M.compare <- compare(M1.loo.out, M2.loo.out, M3.loo.out, 
                     M4.loo.out, M5.loo.out, M6.loo.out) 

print(M.compare)
```

It is interesting that two worst models according to Pareto k statistics are the two best ones according to looic values... M5 seems best based on Pareto k statistics but worst according to looic.  

```{r}
m <- 4
tmp <- data.frame(M.compare) %>%
  select(elpd_diff, looic, se_looic)
knitr::kable(tmp, format = "html", digits = 2,
      caption = "A comparison of five models based on their predictaive performance. ")
```


Here, I use M4 for the predictions. 

```{r posterior_M4_theta, fig.cap="Figure. Posterior distributions of slope parameters for M8"}

bayesplot::mcmc_dens(M4.jm$samples, c("theta[1]", "theta[2]", "theta[3]", "theta[4]",
                                      "theta[5]", "theta[6]", "theta[7]", "theta[8]",
                                      "theta[9]", "theta[10]", "theta[11]", "theta[12]"))
```

```{r posterior_M4_sigmas, fig.cap="Figure. Posterior distributions of process and sampling standard deviations"}

bayesplot::mcmc_dens(M4.jm$samples, c("sigma.pro1", "sigma.obs"))
```



```{r}
Xs.stats <- M4.Xs.stats
ys.stats <- M4.ys.stats
p.1 <- ggplot() +
  geom_ribbon(data = Xs.stats,
              aes(x = time, 
                  ymin = exp(low_X), 
                  ymax = exp(high_X)),
              fill = fill.color,
              alpha = fill.alpha) +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  # geom_line(data = Xs.stats,
  #           aes(x = time, y = exp(high_X)), 
  #           color = "purple",
  #           linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = exp(median_X)), 
             color = line.color,
             size = 1) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(median_X)), 
            color = line.color,
            alpha = 0.5) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), 
             color = obsd.color,
             alpha = 0.5) + 
  scale_x_continuous(breaks = seq(year.begin, year.end, 5),
                     limits = c(year.begin, year.end)) +
  labs(x = '', y = '# nests', title = "Wermon")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

if (save.fig)
  ggsave(p.1, filename = paste0("figures/W_imputed_M", m, ".png"),
         device = "png", height = 4, width = 6, units = "in",
         dpi = 600)

p.1.log <- ggplot() +
  geom_ribbon(data = Xs.stats,
              aes(x = time, 
                  ymin = (low_X), 
                  ymax = (high_X)),
              fill = fill.color,
              alpha = fill.alpha) +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  # geom_line(data = Xs.stats,
  #           aes(x = time, y = exp(high_X)), 
  #           color = "purple",
  #           linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = (median_X)), 
             color = line.color,
             size = 1) +
  geom_line(data = Xs.stats,
            aes(x = time, y = (median_X)), 
            color = line.color,
            alpha = 0.5) +
  geom_point(data = ys.stats,
             aes(x = time, y = log(obsY)), 
             color = obsd.color,
             alpha = 0.5) + 
  scale_x_continuous(breaks = seq(year.begin, year.end, 5),
                     limits = c(year.begin, year.end)) +
  labs(x = '', y = 'log(# nests)', title = "Wermon")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

if (save.fig)
  ggsave(p.1.log, filename = paste0("figures/W_log_imputed_M", m, ".png"),
         device = "png", height = 4, width = 6, units = "in",
         dpi = 600)

print(p.1.log)
```

Combine all counts per year.

```{r }

Xs.stats %>% mutate(season = ifelse(month < 4, year - 1, year)) %>% 
  mutate(summer = ifelse(month > 3 & month < 10, 1, 0)) -> Xs.stats

ys.stats %>% mutate(season = ifelse(month < 4, year - 1, year)) %>% 
  mutate(summer = ifelse(month > 3 & month < 10, 1, 0)) -> ys.stats

seasons <- as.matrix(unique(Xs.stats$season))

#######################
X.posterior.seasons <- lapply(apply(seasons, 
                                    MARGIN = 1,
                                    FUN = extract.posterior.jagsUI, 
                                    Xs.stats = Xs.stats, 
                                    samples = M1.jm$samples),
                              FUN = function(x){
                                n.summer <- exp(x$samples[, x$summer == 1]) %>% rowSums() 
                                n.winter <- exp(x$samples[, x$summer == 0]) %>% rowSums() 
                                n.season <- exp(x$samples) %>% rowSums()
                                return(data.frame(summer = n.summer, 
                                                  winter = n.winter,
                                                  all = n.season))
                              } )


Xs.season <- as.data.frame(matrix(unlist(lapply(X.posterior.seasons, 
                                                FUN = function(x){
                                                  qtiles <- apply(x, 
                                                                  MARGIN = 2,
                                                                  FUN = quantile, 
                                                                  probs = c(0.025, 0.5, 0.975))})),
                                  ncol = 9, byrow = T))

colnames(Xs.season) <- c("Summer.low", "Summer.median", "Summer.high",
                         "Winter.low", "Winter.median", "Winter.high",
                         "all.low", "all.median", "all.high")
Xs.season <- mutate(Xs.season, 
                    season = as.vector(seasons))

ys.stats %>% group_by(season, summer) %>%
  summarize(obs = sum(obsY)) -> tmp.y.season

ys.season <- data.frame(season = seasons,
                        summer = unlist(c(as.vector(filter(tmp.y.season, summer == 1)[, "obs"]))),
                        winter = unlist(filter(tmp.y.season, summer == 0)[, "obs"]))

Xs.season %>% left_join(ys.season, by = "season") -> Xy.season

if (!file.exists(paste0("data/estimatedX_W_M", m, "_", run.date, ".csv")))
  write.csv(Xy.season, 
            file = paste0("data/estimatedX_W_M", m, "_", run.date, ".csv"),
            quote = F, row.names = F)


p.W.estimated.counts <- ggplot() + 
  geom_ribbon(data = Xs.season,
              aes(x = season, 
                  ymin = Summer.low, 
                  ymax = Summer.high),
              fill = fill.color.summer,
              alpha = fill.alpha) +
  geom_point(data = Xs.season,
             aes(x = season, 
                 y = Summer.median), 
             color = line.color.summer,
             size = 2) +
  geom_point(data = ys.season,
             aes(x = season, 
                 y = summer),
             color = data.color,
             size = 2) +
  
  geom_ribbon(data = Xs.season,
                aes(x = season, 
                    ymin = Winter.low, 
                    ymax = Winter.high), 
                fill = fill.color.winter,
              alpha = fill.alpha) + 
  geom_point(data = Xs.season,
             aes(x = season, 
                 y = Winter.median), 
             color = line.color.winter,
             shape = 17,
             size = 2)+
  
  geom_point(data = ys.season,
             aes(x = season, 
                 y = winter),
             color = data.color,
             shape = 17,
             size = 2) + 
  scale_x_continuous(breaks = seq(year.begin, year.end, 5),
                     limits = c(year.begin, year.end)) +
  labs(x = '', y = '# nests', title = "Wermon")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
  
  #labs(title = 'Jamursba-Medi', x = 'Season', y = "Nest counts")

if (save.fig)
  ggsave(filename = "figures/CountsPerSeason_best_W.png",
         device = "png", dpi = 600,height = 4, width = 6, units = "in",
         plot = p.W.estimated.counts)
print(p.W.estimated.counts)
```

