---
title: "Nest estimations"
output: html_notebook
---

This analysis tries to fill in missing nest counts data. A parametric model is fitted to years with complete data. The estimated parameters are used to fill in years without complete data. Ultimately, it aims to look at how the counts are changing over time. 

```{r} 
# load some libraries and functions. 
rm(list=ls())
source('Dc_Indonesia_nesting_fcns.R')
library(lme4)
library(mgcv)
library(rjags)
library(loo)
base_theme <- ggplot2::theme_get()
library(bayesplot)
#
# # set back to the base theme:
ggplot2::theme_set(base_theme)

```


#Jamursba-Medi
```{r}
# get JM data first:
data.0.JM <- read.csv('data/NestCounts_JM_09Feb2018.csv')

# create time-duration filed (in yrs)
# define dates with begin and end dates:
data.0.JM %>% reshape2::melt(id.vars = "YEAR", 
                             variable.name = "month",
                             value.name = "count") -> data.1.JM
data.1.JM$MONTH <- unlist(lapply(data.1.JM$month, FUN = mmm2month))

data.1.JM <- mutate(data.1.JM, f.month = as.factor(MONTH),
                    f.year = as.factor(YEAR))  %>%
  mutate(Frac.Year = YEAR + (MONTH-0.5)/12) %>%
  reshape::sort_df(.,vars = "Frac.Year")

data.1.2005.JM <- filter(data.1.JM, YEAR > 2004)

```

Take a look at the data:

```{r warning=FALSE}
ggplot(data.1.JM) + 
  geom_point(aes(x = MONTH, y = count, color = f.year)) + 
  geom_line(aes(x = MONTH, y = count, color = f.year)) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = 'Month', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
```

Take a look at the time series since 2001. 

```{r warning=FALSE}
#data.1.JM %>% mutate(Frac.Year = YEAR + (MONTH-0.5)/12) -> data.1.JM
data.1.JM.2001 <- filter(data.1.JM, YEAR > 2000)
ggplot(data.1.JM.2001) + 
  geom_point(aes(x = Frac.Year, y = count)) + 
  geom_line(aes(x = Frac.Year, y = count)) +
  scale_x_continuous(breaks = seq(2000, 2018, 5),
                     limits = c(2000, 2018)) +
  labs(x = 'Time', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
```

Using the cyclic nature of nesting, we should be able to fit a model to fill in the missing data points. We need to make some assumptions about how data were collected, unless we have a measure of effort every year.  I ran time-series models to compare how they do in terms of filling in the missing data points. Here is a state-space approach using AR(1) process and using month as the indicator for picking one of two variances in the process. Because it takes a bit of time to run these, output has been saved as an RData file. Bring it in and make some plots here. I load results here.

First, look at the model statement in JAGS language.
```{r}
load("RData/SSAR1_month_JM_2001_2018-04-25.RData")
results.JM_SSAR1_month$jm
```

This dataset starts at 2001 and there are three missing values at the beginning. 

```{r}
results.JM_SSAR1_month$bugs.data
```

The MCMC run converged fine according to the Gelman diagnostic.

```{r}
results.JM_SSAR1_month$g.diag
```

Looks okay although the upper end for sigma.obs may be a bit high. The trace plot for the parameter is below.

```{r}
mcmc_trace(results.JM_SSAR1_month$zm, c("sigma.obs"))
```

For the other good behaving parameters, trace plots are much nicer. 
```{r}
mcmc_trace(results.JM_SSAR1_month$zm, c("mu", "sigma.pro1", "sigma.pro2"))
```

The entire posteior of sigma.obs doesn't look too bad.

```{r}
mcmc_dens(results.JM_SSAR1_month$zm, "sigma.obs")
```

Then, look at the predicted values as a time-series plot. Red is the estimated missing values and green are observed counts. 5 and 95% confidence limits are shown in dashed line. Although I called mode_X, they should have been median_X because they were 50 percentiles... oops. 

```{r warning=F}
Xs.stats <- results.JM_SSAR1_month$Xs.stats
ys.stats <- results.JM_SSAR1_month$ys.stats

ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats,
            aes(x = time, y = high_X), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = mode_X), color = "red",
             alpha = 0.5, size = 2) +
  geom_line(data = Xs.stats,
            aes(x = time, y = mode_X), color = "red",
            alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = low_X), color = "red",
            linetype = 2) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5) +
  labs(title = '', x = '', y = "Nest counts")


```

We can also use two theta parameters for increasing and decreasing sides each year. 

```{r}
load('RData/SSAR1_month_JM_var_theta_2001_2018-04-25.RData')
results.JM_SSAR1_month_var_theta$jm
```

Let's make sure I used the right data.

```{r}
results.JM_SSAR1_month_var_theta$bugs.data
```


```{r warning=F}
Xs.stats <- results.JM_SSAR1_month_var_theta$Xs.stats
ys.stats <- results.JM_SSAR1_month_var_theta$ys.stats

ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats,
            aes(x = time, y = high_X), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = mode_X), color = "red",
             alpha = 0.5, size = 2) +
  geom_line(data = Xs.stats,
            aes(x = time, y = mode_X), color = "red",
            alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = low_X), color = "red",
            linetype = 2) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5) +
  labs(title = '', x = '', y = "Nest counts")

```

Confidence intervals are a bit better for the two-theta model than the one-theta model.  So, let's assume this was sufficient...  

```{r echo=FALSE}
results.JM_SSAR1_month_var_theta$g.diag
```

Convergence looks good and posteriors of standard deviations for the process (sigma.pro1 and sigma.pro2) look okay. sigma.pro1 is for May, June, July, and August. 
```{r}
zm <- results.JM_SSAR1_month_var_theta$zm

mcmc_dens(zm, c("sigma.pro1", "sigma.pro2"))
```

```{r}
mcmc_dens(zm, c("theta.1", "theta.2"))

```

theta.1 is for January through July. 

Posteriors of those parameters look okay.  How about sigma.obs?

```{r}
mcmc_dens(zm, "sigma.obs")
```

Not great but not horrible either. Just to double check the convergence visually, look at the trace of these chains.

```{r}
mcmc_trace(zm, "sigma.obs")
```

Not so good... but what can we do to improve? ...

Sum within each year to find change in trend over the years.
```{r}
Xs.stats %>% #mutate(f.year = as.factor(year)) %>%
  group_by(year) %>% summarise(total = sum(mode_X), 
                               low = sum(low_X),
                               high = sum(high_X)) -> Xs.year

ggplot(Xs.year) + 
  geom_point(aes(x = year, y = total))+
  geom_errorbar(aes(x = year, ymin = low, ymax = high)) + 
  labs(title = '', x = '', y = "Nest counts")
```

Unfortunately, confidence intervals are pretty wide for the last few years because of the missing data points... Adding some older data points looks like this. These data points are from Manjula.

```{r}
Manj.data <- data.frame(year = c(1984, 1985, 1993, 1994, 1995,
                                1996, 1997, 1999,2000, 2001,
                                2002, 2003, 2004, 2005, 2006,
                                2007, 2008, 2009, 2010, 2011,
                                2012, 2013, 2014, 2015, 2016),
                       total =  c(14522, 3261, 4448, 4517, 4633,
                                  6929, 4879, 3429, 2383, 3321,
                                  2143, 3790, 3597, 2626, 2674,
                                  2107, 2077, 2055, 1720, 1596,
                                  1257, 473, 902, 836, 1151),
                       low = NA, high = NA)
#all.data <- rbind(old.data, Xs.year)

ggplot() + 
  geom_point(data = Xs.year,
             aes(x = year, y = total))+
  geom_errorbar(data = Xs.year,
                aes(x = year, ymin = low, ymax = high)) + 
  geom_point(data = Manj.data,
             aes(x = year, y = total),
             color = 'red') + 
  labs(title = '', x = '', y = "Nest counts")

```


In the past, Tapilatu reported the declining trend (at >5%/year) using the data from 1984 to 2011. We repeat the same analysis here.

```{r}
dataTo2011 <- Manj.data %>% filter(year < 2012) %>% mutate(year.0 = year - 1984)

fitTo2011 <- glm(log(total) ~ year.0, 
                 family = 'gaussian', 
                 data = dataTo2011)
summary(fitTo2011)

```

Identical result to the paper for JM.

Now create a similar plot here:

```{r}
new.data.JM <- data.frame(year.0 = dataTo2011$year.0)
pred.JM <- predict(fitTo2011, 
                   newdata = new.data.JM, 
                   se.fit = T)

pred.JM.df <- data.frame(year = dataTo2011$year,
                         total = exp(pred.JM$fit),
                         se2.high = exp(pred.JM$fit + pred.JM$se.fit * 2),
                         se2.low = exp(pred.JM$fit - pred.JM$se.fit * 2))

ggplot() + 
  geom_point(data = Xs.year,
             aes(x = year, y = total))+
  geom_errorbar(data = Xs.year,
                aes(x = year, ymin = low, ymax = high)) + 
  geom_point(data = Manj.data,
             aes(x = year, y = total),
             color = 'red') + 
  geom_line(data = pred.JM.df,
            aes(x = year, y = total)) +
  geom_line(data = pred.JM.df,
            aes(x = year, y = se2.high)) +
  geom_line(data = pred.JM.df,
            aes(x = year, y = se2.low)) +

  labs(title = '', x = '', y = "Nest counts")

```


If we add a few more data points, the rate of decline will decrease because of the counts continue to decrease until 2013. However, the number of nests seems to increase from that year. So, it may be possible to fit a model that can accommodate an inflection point. 

```{r}
all.data <- rbind(filter(Manj.data, year < 2000), Xs.year) %>% 
  mutate(year.0 = year - min(year))

fit.all.yrs <- lm(log(total) ~ year.0,data = all.data)
summary(fit.all.yrs)

```

As expected, the rate of decline is >6%. But, is that a fair treatment? 

```{r}
new.data.JM <- data.frame(year.0 = all.data$year.0)

pred.JM.all <- predict(fit.all.yrs, 
                   newdata = new.data.JM, 
                   se.fit = T)

pred.JM.all.df <- data.frame(year = all.data$year,
                         total = exp(pred.JM.all$fit),
                         se2.high = exp(pred.JM.all$fit + pred.JM.all$se.fit * 2),
                         se2.low = exp(pred.JM.all$fit - pred.JM.all$se.fit * 2))

ggplot() + 
  geom_point(data = all.data,
             aes(x = year, y = total))+
  geom_errorbar(data = all.data,
                aes(x = year, ymin = low, ymax = high)) + 
  geom_point(data = Manj.data,
             aes(x = year, y = total),
             color = 'red') + 
  geom_line(data = pred.JM.all.df,
            aes(x = year, y = total)) +
  geom_line(data = pred.JM.all.df,
            aes(x = year, y = se2.high)) +
  geom_line(data = pred.JM.all.df,
            aes(x = year, y = se2.low)) +

  labs(title = '', x = '', y = "Nest counts")
```

What does a GAM say?

```{r}

fit.gam.JM <- gam(total ~ s(year.0), 
                 data = all.data)
sum.gam.JM <- summary(fit.gam.JM)
sum.gam.JM

```

Estimated degrees of freedom (edf) is about 1, indicating a linear fit may be sufficient. But... what "sufficient" is not very interesting. Take a look at the GAM fit.

```{r}
gam.val.JM <- plot(fit.gam.JM)
gam.val.JM.df <- data.frame(x = gam.val.JM[[1]]$x,
                         fit = gam.val.JM[[1]]$fit,
                         se = gam.val.JM[[1]]$se)

all.data %>% mutate(log.total = log(total)) %>%
  mutate(log.total.std = log(total) - mean(log(total), na.rm = T)) %>%
  mutate(total.std = total - mean(total, na.rm = T)) -> all.data

ggplot() + 
  geom_line(data = gam.val.JM.df,
            aes(x = x+1984, y = fit)) + 
  geom_line(data = gam.val.JM.df,
            aes(x = x+1984, y = fit + se), linetype = 2) + 
  geom_line(data = gam.val.JM.df,
            aes(x = x+1984, y = fit - se), linetype = 2) + 
  geom_point(data = all.data,
             aes(x = year, y = total.std),
             color = 'black') + 
  labs(title = '', x = '', 
       y = paste0('GAM(', signif(sum.gam.JM$edf, digits = 3), ')'))
```

Take a look at the data in the log space.
```{r}
ggplot() + 
  geom_point(data = all.data,
             aes(x = year, y = log(total)))+
  geom_errorbar(data = all.data,
                aes(x = year, ymin = log(low), ymax = log(high))) + 
  geom_point(data = Manj.data,
             aes(x = year, y = log(total)),
             color = "red")
```

When we look at this plot long enough, we start to see there is a bit of up-tick after the 30th year (2013). A similar bouncing happened between 2002 and 2004. The decline from 2004 to 2013 is dramatic but how about after that? The uncertainty in this plot is exaggerated because the tim-series model for filling in missing values was not great. (missing values are a serious issue in time-series analysis in general.) Red points indicate Manjula's calculations. 

If we just fit a similar model to the five data points since 2013, this is what we get.
```{r}
dataSince2013 <- filter(all.data, year > 2012)

fit.lm.2013 <- lm(log(total) ~ year.0, data = dataSince2013)
summary(fit.lm.2013)
```

The estimated annual rate of increase is >20%. If we incorporate the uncertainty in these estimates, we can extract the posterior samples and compute the range of slopes.

```{r}
# extract.samples is a function in TomosFnctions.R
X23.2013 <- extract.samples(zm, 'X[23]')
X24.2014 <- extract.samples(zm, 'X[24]')
X25.2015 <- extract.samples(zm, 'X[25]')
X26.2016 <- extract.samples(zm, 'X[26]')
X27.2017 <- extract.samples(zm, 'X[27]')

Xs <- cbind(X23.2013, X24.2014, X25.2015, X26.2016, X27.2017)

index <- seq(1, nrow(Xs), by = nrow(Xs)/1000) 

Xs.samples <- Xs[index,]

fit.all <- apply(Xs.samples, MARGIN = 1, 
                 FUN = function(x) {df <- data.frame(total = x,
                                                     year = 0:4)
                 fit.1 <- lm(log(total) ~ year, data = df)
                 return(fit.1$coefficients[2])})

fit.all.df <- data.frame(slope = fit.all)

p.pos <- length(fit.all.df$slope[fit.all.df$slope > 0])/length(fit.all)
# qtiles <- quantile(fit.all, 
#                    probs = c(0.05, 0.25, 0.5, 0.75, 0.95))

ggplot(data = fit.all.df) + 
  geom_histogram(aes(x = slope),
                 binwidth = 0.1) 
```

So, it is more likely that the slope is positive - BUT the uncertainty is exaggerated so the picture is not as rosy as it seems in this plot. 

#Warmon
```{r}
rm(list=ls())
source('Dc_Indonesia_nesting_fcns.R')

```


```{r}
data.0.W <- read.csv('data/NestCounts_Warmon_27March2018.csv')

# create time-duration filed (in yrs)
# define dates with begin and end dates:
data.0.W %>% reshape2::melt(id.vars = "YEAR", 
                             variable.name = "month",
                             value.name = "count") -> data.1.W
data.1.W$MONTH <- unlist(lapply(data.1.W$month, FUN = mmm2month))

data.1.W <- mutate(data.1.W, f.month = as.factor(MONTH),
                    f.year = as.factor(YEAR))  %>%
  mutate(Frac.Year = YEAR + (MONTH-0.5)/12) %>%
  reshape::sort_df(.,vars = "Frac.Year")

```

Raw data look like this. 
```{r}
ggplot(data.1.W) + 
  geom_point(aes(x = Frac.Year, y = count)) + 
  geom_line(aes(x = Frac.Year, y = count))
```

We see a lot of missing values between the mid 2013 and 2016. That's a tough one to fill in as no data are available at all... Take a look at annual fluctuations:


```{r warning=FALSE}
ggplot(data.1.W) + 
  geom_point(aes(x = MONTH, y = count, color = f.year)) + 
  geom_line(aes(x = MONTH, y = count, color = f.year)) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = 'Month', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

```

There are summer and winter peaks: higher values are found between May and September, and between October and April. 2004 and 2005 seem to be outliers with respect to within-year trend. 

```{r}
data.1.W %>% filter(YEAR > 2005) -> data.1.W.2006
ggplot(data.1.W.2006) + 
  geom_point(aes(x = MONTH, y = count, color = f.year)) + 
  geom_line(aes(x = MONTH, y = count, color = f.year)) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = 'Month', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
```

The pattern looks more prominent between 2006 and 2017 (2014 and 2015 are missing). Low counts happen in March, April, September, and October. 

When converting counts into log(counts);

```{r}
ggplot(data.1.W) + 
  geom_point(aes(x = Frac.Year, y = log(count))) + 
  geom_line(aes(x = Frac.Year, y = log(count)))
```

```{r}
ggplot(data.1.W) + 
  geom_point(aes(x = MONTH, y = log(count), color = f.year)) + 
  geom_line(aes(x = MONTH, y = log(count), color = f.year)) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = 'Month', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
```

Because of the large number of missing data, I fit time-series models up to 2014 (six missing values in 2013). First, fit the entire data up to 2014 using two variance terms; one for March, April, September and October and another for the rest.

```{r}
load("RData/SSAR1_month_Warmon_2018-04-25_To2013.RData")
zm.W <- results.Warmon_SSAR1_month_To2013$zm
Xs.stats.W <- results.Warmon_SSAR1_month_To2013$Xs.stats
ys.stats.W <- results.Warmon_SSAR1_month_To2013$ys.stats
# Gelman-diagnostic seems okay.
#results.Warmon_SSAR1_month_To2013$g.diag
```

The model is the same as before with one theta and two process taus, where tau.pro1 is for March, April, September and October and tau.pro2 is for the rest. These two groups correspond to high and low numbers, respectively. 

```{r}
results.Warmon_SSAR1_month_To2013$jm
```

And data are here:

```{r}
results.Warmon_SSAR1_month_To2013$bugs.data
```


The posterior of sigma.pro1 and sigma.pro2 look like these:

```{r}
mcmc_dens(zm.W, c("sigma.pro1", "sigma.pro2"))
```

And the trace.

```{r}
mcmc_trace(zm.W, c("sigma.pro1", "sigma.pro2"))
```


Then look at the series. 
```{r}
ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats.W,
            aes(x = time, y = high_X), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats.W,
             aes(x = time, y = mode_X), color = "red",
             alpha = 0.5) +
  geom_line(data = Xs.stats.W,
            aes(x = time, y = mode_X), color = "red",
            alpha = 0.5) +
  geom_point(data = ys.stats.W,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5)+
  geom_line(data = ys.stats.W,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5) + 
  labs(x = '', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
```

Seemed to fit fairly well. Now, take out 2004 and 2005 and see how that fits.

```{r}
load("RData/SSAR1_month_Warmon_2018-04-25_2006To2013.RData")
zm.W2 <- results.Warmon_SSAR1_month_2006To2013$zm
Xs.stats.W2 <- results.Warmon_SSAR1_month_2006To2013$Xs.stats
ys.stats.W2 <- results.Warmon_SSAR1_month_2006To2013$ys.stats
```

Make sure the model is still the same.

```{r}
results.Warmon_SSAR1_month_2006To2013$jm
```

And the data:
```{r}
results.Warmon_SSAR1_month_2006To2013$bugs.data
```

Gelman diagnostic:

```{r}
results.Warmon_SSAR1_month_2006To2013$g.diag
```

And the posteriors of process standard deviations.

```{r}
mcmc_dens(zm.W2, c("sigma.pro1", "sigma.pro2"))
```

sigma.pro1, SD for the low count months (March, April, September, and October), is smaller than when the entire data were used. The posterior for the SD of observation model looks okay.  

```{r}
mcmc_dens(zm.W2, "sigma.obs")
```

Predictions series look like this.

```{r}
ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats.W2,
            aes(x = time, y = high_X), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats.W2,
             aes(x = time, y = mode_X), color = "red",
             alpha = 0.5) +
  geom_line(data = Xs.stats.W2,
            aes(x = time, y = mode_X), color = "red",
            alpha = 0.5) +
  geom_point(data = ys.stats.W2,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5)+
  geom_line(data = ys.stats.W2,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5) + 
  labs(x = '', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
```

Now, how about including the two and a half years of missing data and more recent data? 

```{r}
load("RData/SSAR1_month_Warmon_2018-04-25_2006To2017.RData")
zm.W3 <- results.Warmon_SSAR1_month_2006To2017$zm
Xs.stats.W3 <- results.Warmon_SSAR1_month_2006To2017$Xs.stats
ys.stats.W3 <- results.Warmon_SSAR1_month_2006To2017$ys.stats
```

Check to see the model is still the same:

```{r}
results.Warmon_SSAR1_month_2006To2017$jm
```

And the data:
```{r}
results.Warmon_SSAR1_month_2006To2017$bugs.data
```


Then the convergence
```{r}
results.Warmon_SSAR1_month_2006To2017$g.diag
```

Posteriors on process standard deviations.

```{r}
mcmc_dens(zm.W3, c("sigma.pro1", "sigma.pro2"))
```

The posterior of the SD of observation model seems okay.

```{r}
mcmc_dens(zm.W3, "sigma.obs")
```

Then the prediction series:

```{r}
ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats.W3,
            aes(x = time, y = high_X), color = "red",
            linetype = 2) +
  geom_line(data = Xs.stats.W3,
            aes(x = time, y = low_X), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats.W3,
             aes(x = time, y = mode_X), color = "red",
             alpha = 0.5) +
  geom_line(data = Xs.stats.W3,
            aes(x = time, y = mode_X), color = "red",
            alpha = 0.5) +
  geom_point(data = ys.stats.W3,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5)+
  geom_line(data = ys.stats.W3,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5) +
  labs(x = '', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
```

Probably not useful to fill in those years as we have no idea what was happening. For other years, however, the estimated true counts (red) seem reasonable. I will use all years except 2014 and 2015. 2013 has only six data points. I pool estimated counts annually. I add the first two data points as the median values. 

```{r}
Xs.stats.W3 %>% #mutate(f.year = as.factor(year)) %>%
  group_by(year) %>% summarise(total = sum(mode_X), 
                               low = sum(low_X),
                               high = sum(high_X)) -> Xs.year.W3

Xs.year.W3[Xs.year.W3$year == 2014 | Xs.year.W3$year == 2015, 
        c("total", "low", "high")] <- NA

data.1.W %>% filter(YEAR < 2006) %>% group_by(YEAR) %>%
  summarise(total = sum(count)) %>%
  mutate(low = NA, high = NA) %>%
  rename(year = YEAR) -> Xs.year.2004.2005

Xs.year.W3 <- rbind(Xs.year.2004.2005, Xs.year.W3)

ggplot(Xs.year.W3) + 
  geom_point(aes(x = year, y = total))+
  geom_errorbar(aes(x = year, ymin = low, ymax = high)) + 
  labs(title = '', x = '', y = "Nest counts")

```

Uncertainty is so big that we can't really tell how the numbers are chaning over time. Just looking at the mode:

```{r}
ggplot(Xs.year.W3) + 
  geom_point(aes(x = year, y = total))+
  geom_line(aes(x = year, y = total))+
  labs(title = '', x = '', y = "Nest counts")
```

A simple regression analysis up to 2011:

```{r}
dataTo2011.W <- Xs.year.W3 %>% filter(year < 2012) %>% mutate(year.0 = year - 2004)

fitTo2011.W <- glm(log(total) ~ year.0, 
                 family = 'gaussian', 
                 data = dataTo2011.W)
summary(fitTo2011.W)
```

Up to 2011, the rate of decline is 12.7%/year. 

```{r}
new.data.W <- data.frame(year.0 = dataTo2011.W$year.0)
pred.2011.W <- predict(fitTo2011.W, 
                   newdata = new.data.W, 
                   se.fit = T)

pred.2011.W.df <- data.frame(year = dataTo2011.W$year,
                         total = exp(pred.2011.W$fit),
                         se2.high = exp(pred.2011.W$fit + pred.2011.W$se.fit * 2),
                         se2.low = exp(pred.2011.W$fit - pred.2011.W$se.fit * 2))

ggplot() + 
  geom_point(data = Xs.year.W3,
             aes(x = year, y = total))+
  geom_errorbar(data = Xs.year.W3,
                aes(x = year, ymin = low, ymax = high)) + 
  geom_line(data = pred.2011.W.df,
            aes(x = year, y = total)) +
  geom_line(data = pred.2011.W.df,
            aes(x = year, y = se2.high)) +
  geom_line(data = pred.2011.W.df,
            aes(x = year, y = se2.low)) +

  labs(title = '', x = '', y = "Nest counts")
```


What about up to 2017?

```{r}
dataTo2017.W <- Xs.year.W3 %>% mutate(year.0 = year - 2004)

fitTo2017.W <- glm(log(total) ~ year.0, 
                 family = 'gaussian', 
                 data = dataTo2017.W)
summary(fitTo2017.W)
```

The rate of decline is only about 6%/year now. Of course, that's looking just the median values. 

```{r}
new.data.2017.W <- data.frame(year.0 = dataTo2017.W$year.0)
pred.2017.W <- predict(fitTo2017.W, 
                   newdata = new.data.2017.W, 
                   se.fit = T)

pred.2017.W.df <- data.frame(year = dataTo2017.W$year,
                         total = exp(pred.2017.W$fit),
                         se2.high = exp(pred.2017.W$fit + pred.2017.W$se.fit * 2),
                         se2.low = exp(pred.2017.W$fit - pred.2017.W$se.fit * 2))

ggplot() + 
  geom_point(data = Xs.year.W3,
             aes(x = year, y = total))+
  geom_errorbar(data = Xs.year.W3,
                aes(x = year, ymin = low, ymax = high)) + 
  geom_line(data = pred.2017.W.df,
            aes(x = year, y = total)) +
  geom_line(data = pred.2017.W.df,
            aes(x = year, y = se2.high)) +
  geom_line(data = pred.2017.W.df,
            aes(x = year, y = se2.low)) +

  labs(title = '', x = '', y = "Nest counts")
```

What does a GAM say?

```{r}
fit.gam.W <- gam(total ~ s(year.0), 
                 data = dataTo2017.W)
sum.gam.W <- summary(fit.gam.W)
sum.gam.W
#plot(fit.gam.1)
```

ACtually it's a pretty good fit. 

```{r}
gam.val.W <- plot(fit.gam.W)
gam.val.W.df <- data.frame(x = gam.val.W[[1]]$x,
                         fit = gam.val.W[[1]]$fit,
                         se = gam.val.W[[1]]$se)

Xs.year.W3 %>% mutate(log.total = log(total)) %>%
  mutate(log.total.std = log(total) - mean(log(total), na.rm = T)) %>%
  mutate(total.std = total - mean(total, na.rm = T)) -> Xs.year.W3

ggplot() + 
  geom_line(data = gam.val.W.df,
            aes(x = x+2004, y = fit)) + 
  geom_line(data = gam.val.W.df,
            aes(x = x+2004, y = fit + se), linetype = 2) + 
  geom_line(data = gam.val.W.df,
            aes(x = x+2004, y = fit - se), linetype = 2) + 
  geom_point(data = Xs.year.W3,
             aes(x = year, y = total.std),
             color = 'red') + 
  labs(title = '', x = '', 
       y = paste0('GAM(', signif(sum.gam.W$edf, digits = 3), ')'))
  
```

Perhaps, there is an uptick since 2012? 


