---
title: "Nest count imputation - Jamursba Medi"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
source("Dc_Indonesia_nesting_fcns.R")
run.date <- "2019-03-14"
models.df <- model.names()
year.begin <- 1999
year.end <- 2019

fill.color <-  "darkseagreen"
fill.color.summer <- "darksalmon"
fill.color.winter <- "gray65"
fill.alpha <-  0.65
line.color <-  "darkblue"
line.color.summer <- "red3"
line.color.winter <- "greenyellow"
data.color <- "black"
data.size <- 1.5
obsd.color <- "red2"

obsd.size <- 1
```

This document describes how leatherback turtle nest count data were imputed for Jamursba-Medi. All models were ran separately and results stored in .rds files. 

We use the Bayesian state-space approach, where the mean of the state process is modeled with an auto-regressive model with time lag of one step (AR(1)). 

M[t] = S[t] * X[t-1]

where s[t] is the “slope” parameter between two consecutive counts (transformed into the natural log scale). When s[t] > 1, the counts increase whereas s[t] < 1, they decline. This slope parameter can be constant (s[t] = s for all t), grouped over some time periods (s[t] = s1 for some t, s[t] = s2 otherwise), function of other variables (s[t] = f(k), where k is a vector of one or more variables), or all different. The last option, i.e., s[t] is different for all t, probably is not useful as there would be too many parameters.  We consider the other three options in this study. 

The state is modeled with a normal distribution.

X[t] ~ N(M[t], v[t]),

where v[t] is the variance of X[t]. The variance also can be modeled in a few different scenarios as was in the slope parameter. It can be constant (v[t] = v for all t), grouped over some time periods (v[t] = v1 for some t, v[t] = v2 otherwise), function of other variables (v[t] = f(z), where z is a vector of one or more variables), or all different. Again we consider the three options as before. 

Given the state process, observations are modeled with normal or t distributions. 

Y[t] ~ N(X[t], v.obs)

Y[t] ~ t(X[t], v.obs, df)

where v.obs is the observation variance and df is the degrees-of-freedom parameter of the t distribution. The observation variance is assumed to be constant throughout the dataset. In total 8 models were developed and predictive accuracy was compared to select the most accurate model (Table x).  Model selection process is explained in the subsequent section.  The analysis is conducted in the natural logarithm scale. We define the following:

x[t] = ln(X[t])

m[t] = ln(M[t])

y[t] = ln(Y[t])

s[t] = ln(S[t])

The models were fit to datasets from turtle nesting beaches. Bayesian computations were conducted using jags through jagsUI package in R.  

##Model 1
First model is the simplest; the state space is modeled with NOrmal distribution and observation is modeled also with normal distribution. No seasonal change in variance or slope is considered. Then, the model definitions become the following:

m[t] = s + x[t-1]

x[t] ~ N(m[t], v.pro)

y[t] ~ N(x[t], v.obs)

```{r model1, cache=TRUE, include=TRUE}
#M <- readRDS("RData/SSAR1_logY_norm_norm_JM_1999_2018-10-12.rds")
str.root <- strsplit(strsplit(as.character(models.df[1, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M1.loo.out <- M$jags.out$loo.out$loo.out
M1.jm <- M$jags.out$jm
```

Diagnostics of the expected log predictive density indicated that the model was not appropriate, where all estimated Pareto k statistics (a measure of model fit - TRUE?) were greater than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M1.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M1.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M1.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M1, fig.cap = "Figure. PSIS diagnostic plot for M1"}
plot(M1.loo.out)
```

```{r}
pareto.k <- M1.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))


```

##Model 2
The second model contained one slope and two variances: one for nesting season (May through September) and another for the non-nesting season (October through April). 

m[t] = s + x[t-1] 

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v1 if May <= month < September, v2 otherwise.

y[t] ~ N(x[t], v.obs)

```{r model2, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[2, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M2.loo.out <- M$jags.out$loo.out$loo.out
M2.jm <- M$jags.out$jm

```

Similary to the first model, all estimated Pareto k statistics were greater than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M2.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M2.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M2.jm$Rhat, FUN = max, na.rm = T))), 3)```). We do not consider this model further.

```{r PSIS_M2, fig.cap="Figure. PSIS diagnostic plot for M2"}
plot(M2.loo.out)
```


```{r}
pareto.k <- M2.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 3
The third model has a constant slope and month-specific variances.

m[t] = s + x[t-1]

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v.pro[month], month = 1, 2, ..., 12

y[t] ~ N(x[t], v.obs)

```{r model3, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[3, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_norm_varM_JM_1999_2017_",
#                     run.date, ".rds"))
M3.loo.out <- M$jags.out$loo.out$loo.out
M3.jm <- M$jags.out$jm

```

Majority of estimated Pareto k statistics were greater than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M3.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M3.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M3.jm$Rhat, FUN = max, na.rm = T))), 3)```). We do not consider this model further.

```{r PSIS_M3, fig.cap="Figure. PSIS diagnostic plot for M3"}
plot(M3.loo.out)
```

##Model 4

The fourth model is similar to the first except there were two slope parameters: one for January through July and another for August through December. These corresponded to before and after the peak of each nesting season in August. There is only one variance. The model is as follows:

m[t] = s[t] + x[t-1], s[t] = s1 if Jan <= month < 8, s2 otherwise.

x[t] ~ N(m[t], v.pro)

y[t] ~ N(x[t], v.obs)

```{r model4, cache=T, include=TRUE}

str.root <- strsplit(strsplit(as.character(models.df[4, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_norm_theta_JM_1999_2017_",
#                     run.date, ".rds"))
M4.loo.out <- M$jags.out$loo.out$loo.out
M4.jm <- M$jags.out$jm
```

Diagnostics of the expected log predictive density indicated that the model was considerably better than the previous models. The vast majority of estimated Pareto k parameter was less than 0.7 (```r sum(M4.loo.out$diagnostics$pareto_k < 0.7)``` out of ```r length(M4.loo.out$diagnostics$pareto_k)```). Convergence was reached for all monitored parameters (minimum = ```r signif(min(unlist(lapply(M4.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M4.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M4.jm$Rhat, FUN = max, na.rm = T))), 3)```).  

```{r PSIS_M4, fig.cap="Figure. PSIS diagnostic plot for M4"}
plot(M4.loo.out)
```

The compasion of the four models above (M1 through M4) indicated the slope parameter needs not to be constant, which was obvious from the beginning... Variance, however, does not affect the predictive performance. 

```{r}
pareto.k <- M4.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))



```

##Model 5

The fifth model contained two slopes and two variances: one slope for January through July and another slope for August through December and one variance for nesting season (May through September) and another variance for the non-nesting season (October through April)

m[t] = s[t] + x[t-1], s[t] = s1 if Jan <= month < 8, s2 otherwise.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v1 if May <= month < September, v2 otherwise.

y[t] ~ N(x[t], v.obs)

```{r model5, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[5, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

#M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_norm_var_theta_JM_1999_2017_", run.date, ".rds"))
M5.loo.out <- M$jags.out$loo.out$loo.out
M5.jm <- M$jags.out$jm

```

There were some large estimated Pareto k parameter values, where ```r sum(M5.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M5.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M5.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M5.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M5.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M5, fig.cap="Figure. PSIS diagnostic plot for M5"}
plot(M5.loo.out)
```

```{r}
pareto.k <- M5.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 6

In Model 6, a unique variance parameter was given to each month, whereas two seasonal slopes were defined as before.  
m[t] = s[t] + x[t-1], s[t] = s1 if Jan <= month < 8, s2 otherwise.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v.pro[month], month = 1, ..., 12

y[t] ~ N(x[t], v.obs)

```{r model6, cache=TRUE, include=TRUE}

str.root <- strsplit(strsplit(as.character(models.df[6, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_norm_varM_theta_JM_1999_2017_",
#              run.date, ".rds"))
M6.loo.out <- M$jags.out$loo.out$loo.out
M6.jm <- M$jags.out$jm

```

There were some large estimated Pareto k parameter values, where ```r sum(M6.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M6.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M6.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M6.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M6.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M6, fig.cap="Figure. PSIS diagnostic plot for M6"}
plot(M6.loo.out)
```

So, having a flexibility in variance does not help preditive performance for this dataset. 

```{r}
pareto.k <- M6.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))
```

##Model 7

Model 7, a unique slope parameter was given to each month, whereas there was one variance. 
m[t] = s[t] + x[t-1], s[t] = s[month], where month = 1, ..., 12.

x[t] ~ N(m[t], v.pro), 

y[t] ~ N(x[t], v.obs)

```{r model7, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[7, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

#M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_norm_thetaM_JM_1999_2017_", run.date, ".rds"))
M7.loo.out <- M$jags.out$loo.out$loo.out
M7.jm <- M$jags.out$jm
```

There were some large estimated Pareto k parameter values, where ```r sum(M7.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M7.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M7.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M7.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M7.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M7, fig.cap="Figure. PSIS diagnostic plot for M7"}
plot(M7.loo.out)
```

Having one variance, instead of two or more seems to help the predictive performance.

```{r}
pareto.k <- M7.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 8

In Model 8, a unique slope parameter was given to each month, whereas there were two variance2. 
m[t] = s[t] + x[t-1], s[t] = s[month], where month = 1, ..., 12.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v1 if May <= month < September, v2 otherwise. 

y[t] ~ N(x[t], v.obs, df)

```{r model8, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[8, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

#M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_norm_var_thetaM_JM_1999_2017_", run.date, ".rds"))
M8.loo.out <- M$jags.out$loo.out$loo.out
M8.jm <- M$jags.out$jm

```

There were some large estimated Pareto k parameter values, where ```r sum(M8.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M8.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M8.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M8.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M8.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M8, fig.cap="Figure. PSIS diagnostic plot for M8"}
plot(M8.loo.out)
```

Again, with two variances, the predictive performance is poor.

```{r}
pareto.k <- M8.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 9

In Model 9, a unique slope and variance parameters were given to each month. 
m[t] = s[t] + x[t-1], s[t] = s[month], where month = 1, ..., 12.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v.pro[month], month = 1, ..., 12

y[t] ~ N(x[t], v.obs)

```{r model9, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[9, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

#M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_norm_varM_thetaM_JM_1999_2017_", 
#run.date, ".rds"))

M9.loo.out <- M$jags.out$loo.out$loo.out
M9.jm <- M$jags.out$jm

```

There were some large estimated Pareto k parameter values, where ```r sum(M9.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M9.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M9.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M9.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M9.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M9, fig.cap="Figure. PSIS diagnostic plot for M9"}
plot(M9.loo.out)
```

```{r}
pareto.k <- M9.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 10

Model 10 is similar to the first model but uses a different distribution for the state model; the observation is  modeled with t distribution. No seasonal change in variance or slope is considered. 

m[t] = s + x[t-1]

x[t] ~ N(m[t], v.pro)

y[t] ~ t(x[t], v.obs, df)

```{r model10, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[10, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_JM_1999_2017_",
#                     run.date, ".rds"))
M10.loo.out <- M$jags.out$loo.out$loo.out
M10.jm <- M$jags.out$jm
```

There were some large estimated Pareto k parameter values, where ```r sum(M10.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M10.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M10.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M10.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M10.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M10, fig.cap="Figure. PSIS diagnostic plot for M10"}
plot(M10.loo.out)
```

```{r}
pareto.k <- M10.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red")) 

```

##Model 11

The next model contained one slope and two variances: one for nesting season (May through September) and another for the non-nesting season (October through April) and t distribution for observations. 

m[t] = s + x[t-1] 

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v1 if May <= month < September, v2 otherwise.

y[t] ~ t(x[t], v.obs, df)

```{r model11, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[11, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_var_JM_1999_2017_",
#                     run.date, ".rds"))
M11.loo.out <- M$jags.out$loo.out$loo.out
M11.jm <- M$jags.out$jm

```

Similary to the first model, all estimated Pareto k statistics were greater than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M11.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M11.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M11.jm$Rhat, FUN = max, na.rm = T))), 3)```). We do not consider this model further.

```{r PSIS_M11, fig.cap="Figure. PSIS diagnostic plot for M11."} 
plot(M11.loo.out)
```

```{r}
pareto.k <- M11.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 12

Model 12 has a constant slope and month-specific variances. The model is as follows:

m[t] = s + x[t-1] 

x[t] ~ N(m[t], v.pro[t]), where v.pro[t] = v.pro[month], month = 1, 2, ..., 12

y[t] ~ t(x[t], v.obs, df)


```{r model12, cache=T, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[12, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_varM_JM_1999_2017_", 
#                     run.date, ".rds"))
M12.loo.out <- M$jags.out$loo.out$loo.out
M12.jm <- M$jags.out$jm
```

The vast majority of estimated Pareto k parameter was greater than 0.7 (```r sum(M12.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M12.loo.out$diagnostics$pareto_k)```. Convergence was reached for all monitored parameters (minimum = ```r signif(min(unlist(lapply(M12.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M12.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M12.jm$Rhat, FUN = max, na.rm = T))), 3)```).  

```{r PSIS_M12, fig.cap="Figure. PSIS diagnostic plot for M12"}
plot(M12.loo.out)
```

```{r}
pareto.k <- M12.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red")) 

```

##Model 13

Model 13 is similar to the second except observations are modeled with t distribution.  These corresponded to before and after the peak of each nesting season in August. The model is as follows:

m[t] = s[t] + x[t-1], s[t] = s1 if Jan <= month < 8, s2 otherwise.

x[t] ~ N(m[t], v.pro)

y[t] ~ t(x[t], v.obs, df)


```{r model13, cache=T, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[13, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_theta_JM_1999_2017_", 
#                     run.date, ".rds"))
M13.loo.out <- M$jags.out$loo.out$loo.out
M13.jm <- M$jags.out$jm
```

All estimated Pareto k parameter was less than 0.7 (```r sum(M13.loo.out$diagnostics$pareto_k < 0.7)``` out of ```r length(M13.loo.out$diagnostics$pareto_k)```. Convergence was reached for all monitored parameters (minimum = ```r signif(min(unlist(lapply(M13.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M13.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M13.jm$Rhat, FUN = max, na.rm = T))), 3)```).  

```{r PSIS_M13, fig.cap="Figure. PSIS diagnostic plot for M13"}
plot(M13.loo.out)
```

```{r}
pareto.k <- M13.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red")) 

```

##Model 14

Next model contained two slopes and two variances: one slope for January through July and another slope for August through December and one variance for nesting season (May through September) and another variance for the non-nesting season (October through April). Observations are modeled with t distribution.

m[t] = s[t] + x[t-1], s[t] = s1 if Jan <= month < 8, s2 otherwise.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v1 if May <= month < September, v2 otherwise.

y[t] ~ N(x[t], v.obs, df)

```{r model14, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[14, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_var_theta_JM_1999_2017_", 
#                     run.date, ".rds"))
M14.loo.out <- M$jags.out$loo.out$loo.out
M14.jm <- M$jags.out$jm

```

There were some large estimated Pareto k parameter values, where ```r sum(M14.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M14.loo.out$diagnostics$pareto_k)``` were larger than 0.7.  Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M14.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M14.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M14.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M14, fig.cap="Figure. PSIS diagnostic plot for M14."}
plot(M14.loo.out)
```

```{r}
pareto.k <- M14.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 15

In Model 15, a unique variance parameter was given to each month, whereas two seasonal slopes were defined as before.  
m[t] = s[t] + x[t-1], s[t] = s1 if Jan <= month < 8, s2 otherwise.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v.pro[month], month = 1, ..., 12

y[t] ~ t(x[t], v.obs, df)

```{r model15, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[15, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_varM_theta_JM_1999_2017_", 
#                     run.date, ".rds"))
M15.loo.out <- M$jags.out$loo.out$loo.out
M15.jm <- M$jags.out$jm

```

Almost all estimated Pareto k parameter values were greater than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M15.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M15.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M15.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M15, fig.cap="Figure. PSIS diagnostic plot for M15."}
plot(M15.loo.out)
```

```{r}
pareto.k <- M15.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 16

In Model 16, a unique slope parameter was given to each month, whereas there was one variance. 
m[t] = s[t] + x[t-1], s[t] = s[month], where month = 1, ..., 12.

x[t] ~ N(m[t], v.pro), 

y[t] ~ t(x[t], v.obs, df)

```{r model16, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[16, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_thetaM_JM_1999_2017_",
#                     run.date, ".rds"))
M16.loo.out <- M$jags.out$loo.out$loo.out
M16.jm <- M$jags.out$jm

```

There were some large estimated Pareto k parameter values, where ```r sum(M16.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M16.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M16.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M16.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M16.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M16, fig.cap="Figure. PSIS diagnostic plot for M16."}
plot(M16.loo.out)
```

```{r}
pareto.k <- M16.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red")) 

```

##Model 17

In Model 17, a unique slope parameter was given to each month, whereas there were two variances. 
m[t] = s[t] + x[t-1], s[t] = s[month], where month = 1, ..., 12.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v1 if May <= month < September, v2 otherwise.

y[t] ~ t(x[t], v.obs, df)

```{r model17, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[17, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_var_thetaM_JM_1999_2017_",
#                     run.date, ".rds"))
M17.loo.out <- M$jags.out$loo.out$loo.out
M17.jm <- M$jags.out$jm

```

There were many large estimated Pareto k parameter values, where ```r sum(M17.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M17.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M17.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M17.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M17.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M17, fig.cap="Figure. PSIS diagnostic plot for M17."}
plot(M17.loo.out)
```

```{r}
pareto.k <- M17.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red"))

```

##Model 18

In Model 18, a unique slope and variance parameters were given to each month. 
m[t] = s[t] + x[t-1], s[t] = s[month], where month = 1, ..., 12.

x[t] ~ N(m[t], v.pro[t]), v.pro[t] = v.pro[month], month = 1, ..., 12

y[t] ~ t(x[t], v.obs, df)

```{r model18, cache=TRUE, include=TRUE}
str.root <- strsplit(strsplit(as.character(models.df[18, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

# M <- readRDS(paste0("RData/jagsout_SSAR1_logY_norm_t_varM_thetaM_JM_1999_2017_",
#                     run.date, ".rds"))
M18.loo.out <- M$jags.out$loo.out$loo.out
M18.jm <- M$jags.out$jm

```

There were some large estimated Pareto k parameter values, where ```r sum(M18.loo.out$diagnostics$pareto_k > 0.7)``` out of ```r length(M18.loo.out$diagnostics$pareto_k)``` were larger than 0.7. Convergence statistics indicated acceptable convergence among the monitored parameters (minimum = ```r signif(min(unlist(lapply(M18.jm$Rhat, FUN = min, na.rm = T))), 3)```, mean = ```r signif(min(unlist(lapply(M18.jm$Rhat, FUN = mean, na.rm = T))), 3)```, and maximum = ```r signif(min(unlist(lapply(M18.jm$Rhat, FUN = max, na.rm = T))), 3)```). 

```{r PSIS_M18, fig.cap="Figure. PSIS diagnostic plot for M18."}
plot(M18.loo.out)
```

```{r}
pareto.k <- M18.loo.out$diagnostics$pareto_k
data.y <- na.omit(M$jags.out$jags.data$y)

pareto.df <- data.frame(y = data.y,
                        khat = pareto.k,
                        datapoint = seq(from = 1, to = length(data.y)),
                        k0.7 = cut(pareto.k,
                                   breaks = c(0, 0.7, 1.5),
                                   labels = c("<=0.7", ">0.7")))
ggplot(data = pareto.df) +   
  geom_path(aes(x = datapoint, y = exp(y)), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = exp(y), 
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+ 
  scale_color_manual(values = c("<=0.7" = "black", 
                                ">0.7" = "red")) 

```

#Model Comparison
First, we compared all models using using ELPD values although Pareto k statistics indicated that models 2, 8, 11, 13, and 17 had acceptable predictive performance. 

```{r ModelComparison, include=TRUE}
# M.compare <- compare(M1.loo.out, M2.loo.out, M3.loo.out, 
#                      M4.loo.out, M5.loo.out, M6.loo.out, 
#                      M7.loo.out, M8.loo.out, M9.loo.out, 
#                      M10.loo.out, M11.loo.out, M12.loo.out, 
#                      M13.loo.out, M14.loo.out, M15.loo.out,
#                      M16.loo.out, M17.loo.out, M18.loo.out)

M.compare <- compare(M2.loo.out, M8.loo.out, M11.loo.out, 
                     M13.loo.out, M17.loo.out) 

print(M.compare)
```

```{r}
tmp <- data.frame(M.compare) %>%
  select(elpd_diff, looic, se_looic)
knitr::kable(tmp, format = "html", digits = 2,
      caption = "A comparison of five models based on their predictaive performance. ")
```


When comparing all models, it was interesting to see that those models that were acceptable according to Pareto k statistics are worse ones according to looic scores. Here, I use M8 and M17 for the predictions. 

Model 8, 

```{r posterior_M8_theta, fig.cap="Figure. Posterior distributions of slope parameters for M8"}
m <- 8
str.root <- strsplit(strsplit(as.character(models.df[m, 'names']), "model_")[[1]][2], ".txt")[[1]]
M8 <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))
M8.jm <- M8$jags.out$jm
bayesplot::mcmc_dens(M8.jm$samples, c("theta.1[1]", "theta.1[2]", "theta.1[3]", "theta.1[4]",
                                      "theta.1[5]", "theta.1[6]", "theta.1[7]", "theta.1[8]",
                                      "theta.1[9]", "theta.1[10]", "theta.1[11]", "theta.1[12]"))
```

```{r posterior_M8_sigmas, fig.cap="Figure. Posterior distributions of process and sampling standard deviations"}

bayesplot::mcmc_dens(M8.jm$samples, c("sigma.pro1", "sigma.obs"))
```


Model 17
```{r posterior_M17_theta, fig.cap="Figure. Posterior distributions of slope parameters for M8"}
m <- 17
str.root <- strsplit(strsplit(as.character(models.df[m, 'names']), "model_")[[1]][2], ".txt")[[1]]
M17 <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))
M17.jm <- M17$jags.out$jm
bayesplot::mcmc_dens(M17.jm$samples, c("theta.1[1]", "theta.1[2]", "theta.1[3]", "theta.1[4]",
                                      "theta.1[5]", "theta.1[6]", "theta.1[7]", "theta.1[8]",
                                      "theta.1[9]", "theta.1[10]", "theta.1[11]", "theta.1[12]"))
```

```{r posterior_M17_sigmas, fig.cap="Figure. Posterior distributions of process and sampling standard deviations"}

bayesplot::mcmc_dens(M17.jm$samples, c("sigma.pro1", "sigma.obs"))
```

```{r posterior_M17_df, fig.cap="Figure. Posterior distributions of degrees-of-freedom"}

bayesplot::mcmc_dens(M17.jm$samples, c("df"))
```

```{r}
Xs.stats <- M8$Xs.stats
ys.stats <- M8$ys.stats
p.1 <- ggplot() +
  geom_ribbon(data = Xs.stats,
              aes(x = time, 
                  ymin = exp(low_X), 
                  ymax = exp(high_X)),
              fill = fill.color,
              alpha = fill.alpha) +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  # geom_line(data = Xs.stats,
  #           aes(x = time, y = exp(high_X)), 
  #           color = "purple",
  #           linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = exp(median_X)), 
             color = line.color,
             size = 1) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(median_X)), 
            color = line.color,
            alpha = 0.5) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), 
             color = obsd.color,
             alpha = 0.5) + 
  scale_x_continuous(breaks = seq(1999, 2019, 5),
                     limits = c(1999, 2019)) +
  labs(x = '', y = '# nests', title = "Jamursba-Medi")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

ggsave(p.1, filename = "figures/JM_imputed_M8.png",
       device = "png", height = 4, width = 6, units = "in",
       dpi = 600)

print(p.1)
```

```{r}

Xs.stats <- M17$Xs.stats
ys.stats <- M17$ys.stats
p.1 <- ggplot() +
  geom_ribbon(data = Xs.stats,
              aes(x = time, 
                  ymin = exp(low_X), 
                  ymax = exp(high_X)),
              fill = fill.color,
              alpha = fill.alpha) +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  # geom_line(data = Xs.stats,
  #           aes(x = time, y = exp(high_X)), 
  #           color = "purple",
  #           linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = exp(median_X)), 
             color = line.color,
             alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(median_X)), 
            color = line.color,
            alpha = 0.5) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), 
             color = obsd.color,
             alpha = 0.5) + 
  scale_x_continuous(breaks = seq(1999, 2019, 5),
                     limits = c(1999, 2019)) +
  labs(x = '', y = '# nests', title = "Jamursba-Medi")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

print(p.1)
```

They are about the same so go with a simpler one: M8. 

Combine all counts per year.

```{r }
m <- 8#5 #2 

str.root <- strsplit(strsplit(as.character(models.df[m, 'names']), "model_")[[1]][2], ".txt")[[1]]
M <- readRDS(paste0("RData/jagsout_", str.root, "_JM_", year.begin, "_", year.end, "_",
                    run.date, ".rds"))

M$Xs.stats %>% mutate(season = ifelse(month < 4, year - 1, year)) %>% 
  mutate(summer = ifelse(month > 3 & month < 10, 1, 0)) -> Xs.stats

M$ys.stats %>% mutate(season = ifelse(month < 4, year - 1, year)) %>% 
  mutate(summer = ifelse(month > 3 & month < 10, 1, 0)) -> ys.stats

seasons <- as.matrix(unique(Xs.stats$season))

#######################
X.posterior.seasons <- lapply(apply(seasons, 
                                    MARGIN = 1,
                                    FUN = extract.posterior.jagsUI, 
                                    Xs.stats = Xs.stats, 
                                    samples = M$jags.out$jm$samples),
                              FUN = function(x){
                                n.summer <- exp(x$samples[, x$summer == 1]) %>% rowSums() 
                                n.winter <- exp(x$samples[, x$summer == 0]) %>% rowSums() 
                                n.season <- exp(x$samples) %>% rowSums()
                                return(data.frame(summer = n.summer, 
                                                  winter = n.winter,
                                                  all = n.season))
                              } )


Xs.season <- as.data.frame(matrix(unlist(lapply(X.posterior.seasons, 
                                                FUN = function(x){
                                                  qtiles <- apply(x, 
                                                                  MARGIN = 2,
                                                                  FUN = quantile, 
                                                                  probs = c(0.025, 0.5, 0.975))})),
                                  ncol = 9, byrow = T))

colnames(Xs.season) <- c("Summer.low", "Summer.median", "Summer.high",
                         "Winter.low", "Winter.median", "Winter.high",
                         "all.low", "all.median", "all.high")
Xs.season <- mutate(Xs.season, 
                    season = as.vector(seasons))

ys.stats %>% group_by(season, summer) %>%
  summarize(obs = sum(obsY)) -> tmp.y.season

ys.season <- data.frame(season = seasons,
                        summer = unlist(c(NA, as.vector(filter(tmp.y.season, summer == 1)[, "obs"]))),
                        winter = unlist(filter(tmp.y.season, summer == 0)[, "obs"]))

Xs.season %>% left_join(ys.season, by = "season") -> Xy.season

if (!file.exists(paste0("data/estimatedX_JM_M", m, "_", run.date, ".csv")))
  write.csv(Xy.season, 
            file = paste0("data/estimatedX_JM_M", m, "_", run.date, ".csv"),
            quote = F, row.names = F)


p.JM.estimated.counts <- ggplot() + 
  geom_ribbon(data = Xs.season,
              aes(x = season, 
                  ymin = Summer.low, 
                  ymax = Summer.high),
              fill = fill.color.summer,
              alpha = fill.alpha) +
  geom_point(data = Xs.season,
             aes(x = season, 
                 y = Summer.median), 
             color = line.color.summer) +
  geom_point(data = ys.season,
             aes(x = season, 
                 y = summer),
             color = data.color) +
  
  geom_ribbon(data = Xs.season,
                aes(x = season, 
                    ymin = Winter.low, 
                    ymax = Winter.high), 
                fill = fill.color.winter,
              alpha = fill.alpha) + 
  geom_point(data = Xs.season,
             aes(x = season, 
                 y = Winter.median), 
             color = line.color.winter)+
  
  geom_point(data = ys.season,
             aes(x = season, 
                 y = winter),
             color = data.color,
             shape = 17) + 
  scale_x_continuous(breaks = seq(1999, 2019, 5),
                     limits = c(1999, 2019)) +
  labs(x = '', y = '# nests', title = "Jamursba-Medi")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
  
  #labs(title = 'Jamursba-Medi', x = 'Season', y = "Nest counts")


ggsave(filename = "figures/CountsPerSeason_best_JM.png",
       device = "png", dpi = 600,height = 4, width = 6, units = "in",
       plot = p.JM.estimated.counts)
print(p.JM.estimated.counts)
```

