---
title: "Combining data imputation and trend analysis using discrete Fourier series with constant CV for two Qs and single U"
output: html_notebook
---

Trying to merge the data imputation process and trend analysis with multiple datasets.

Set things up:

```{r}
rm(list=ls())

# tic <- Sys.time()
# Sys <- Sys.info()
source('Dc_Indonesia_nesting_fcns.R')
#library(rjags)

library(jagsUI)
library(coda)
library(tidyverse)
library(loo)

save.data <- T
save.fig <- F
run.date <- Sys.Date()

```

Set up the plotting colors.
```{r}

fill.color <-  "darkseagreen"
fill.color.N1 <- "blue4"
fill.color.N2 <- "gold4"
fill.color.summer <- "darksalmon"
fill.color.winter <- "gray65"
fill.alpha <-  0.65
line.color <-  "darkblue"
line.color.N <- "cadetblue"
line.color.summer <- "red3"
line.color.winter <- "greenyellow"
data.color <- "black"
data.size <- 1.5
obsd.color <- "red2"

```



Set up the MCMC parameters

```{r}
MCMC.n.chains <- 5
MCMC.n.samples <- 50000
MCMC.n.burnin <- 10000
MCMC.n.thin <- 4

MCMC.params <- list(n.chains = MCMC.n.chains,
                    n.samples = MCMC.n.samples,
                    n.burnin = MCMC.n.burnin,
                    n.thin = MCMC.n.thin)
```


Bring in the data

```{r}
# JM
year.begin <- 2003
year.end <- 2018
# loc <- "JM"
# loc.name <- "Jamursba-Medi"
period.JM <- 12
maxN <- 10000

# W
period.W <- 6

data.jags.JM <- data.extract(location = "JM", 
                             year.begin = year.begin, 
                             year.end = year.end)

data.jags.W <- data.extract(location = "W", 
                             year.begin = year.begin, 
                             year.end = year.end)

filename.root <- paste0("SSAR1_norm_norm_trend_Four_singleU_constCV_independentUQs_",
                        "JM", period.JM, "_W", period.W, "_",
                        year.begin, "_", year.end, "_", run.date)
```

Combine datasets for analysis

```{r}
y <- array(data= NA, dim = c(nrow(data.jags.JM$jags.data2$y),
                             ncol(data.jags.JM$jags.data2$y), 2))
y[,,1] <- data.jags.JM$jags.data2$y
y[,,2] <- data.jags.W$jags.data2$y

jags.data <- list(y = y,
                  #m = data.jags.JM$jags.data2$m[1,],
                  C0 = c(15, 15),
                  n.months = 12,
                  C_cos = c(sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.JM))), 
                            sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.W)))),
                  C_sin = c(sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.JM))),
                            sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.W)))),
                  pi = pi,
                  period = c(period.JM, period.W),
                  N0_mean = c(log(sum(exp(y[1,,1]), na.rm = T)),
                              log(sum(exp(y[1,,2]), na.rm = T))),
                  N0_sd = c(10, 10))

n.timeseries <- dim(y)[3]

# Model-specific parameters
whichPop <- 1:n.timeseries         # multiple time series -> unique population processes
n.states <- max(whichPop)

# Z <- matrix(0,n.timeseries+1,n.states+1)   # matrix with rows as n.timeseries and cols as n.states (pops)
# Z[n.timeseries+1, ] <- NA                  # add a row of NAs to keep jagsUI from converting single time series matrix into vector
# Z[ , n.states+1] <- NA                     # add a col of NAs to keep jagsUI from converting single state matrix into vector
# for(i in 1:length(whichPop)) Z[i,whichPop[i]] <- 1
# 
# jags.data$Z <- Z
#jags.data$n.timeseries <- n.timeseries
jags.data$n.states <- n.states
jags.data$n.years <- dim(y)[1]

# jags.data$q_alpha <- 2
# jags.data$q_beta <- 0.5
# jags.data$r_alpha <- 2
# jags.data$r_beta <- 0.5

```

Define which parameters to monitor

```{r}
jags.params <- c('N', 'U', "p", "p.beta.cos", "p.beta.sin",
                 "sigma.N", 'cv.Q', "sigma.R",
                 "mu", "y", "X", "deviance", "loglik")

```


Run jags

```{r}
# when running with parallel=T, error returns...:
# Error in mcmc.list(x) : Different start, end or thin values in each chain
# Restarting the computer fixed that problem next day... strange...

# it runs fine with no-parallelized - slow but works. 
jm <- jags(jags.data,
           inits = NULL,
           parameters.to.save= jags.params,
           model.file = 'models/model_norm_norm_trend_Four_singleU_constCV_independentQs.txt',
           n.chains = MCMC.params$n.chains,
           n.burnin = MCMC.params$n.burnin,
           n.thin = MCMC.params$n.thin,
           n.iter = MCMC.params$n.samples,
           DIC = T, parallel=T)

#saveRDS(jm, file = paste0("RData/", filename.root, ".rds"))
```


pull together results
```{r}
# extract ys - include estimated missing data
# these need to be arranged in vectors
ys.stats.JM <- data.frame(low = as.vector(t(jm$q2.5$y[,,1])),
                          median = as.vector(t(jm$q50$y[,,1])),
                          high = as.vector(t(jm$q97.5$y[,,1])))
ys.stats.JM$time <- data.jags.JM$data.1$Frac.Year
ys.stats.JM$obsY <- data.jags.JM$data.1$Nests
ys.stats.JM$month <- data.jags.JM$data.1$Month
ys.stats.JM$year <- data.jags.JM$data.1$Year
ys.stats.JM$Season <- data.jags.JM$data.1$Season
ys.stats.JM$location <- "Jamursba-Medi"

ys.stats.W <- data.frame(low = as.vector(t(jm$q2.5$y[,,2])),
                         median = as.vector(t(jm$q50$y[,,2])),
                         high = as.vector(t(jm$q97.5$y[,,2])))
ys.stats.W$time <- data.jags.W$data.1$Frac.Year
ys.stats.W$obsY <- data.jags.W$data.1$Nests
ys.stats.W$month <- data.jags.W$data.1$Month
ys.stats.W$year <- data.jags.W$data.1$Year
ys.stats.W$Season <- data.jags.W$data.1$Season
ys.stats.W$location <- "Wermon"

ys.stats <- rbind(ys.stats.JM, ys.stats.W)

# extract Xs - the state model

Xs.stats.JM <- data.frame(low = as.vector(t(jm$q2.5$X[,,1])),
                          median = as.vector(t(jm$q50$X[,,1])),
                          high = as.vector(t(jm$q97.5$X[,,1])))
Xs.stats.JM$time <- data.jags.JM$data.1$Frac.Year
Xs.stats.JM$obsY <- data.jags.JM$data.1$Nests
Xs.stats.JM$month <- data.jags.JM$data.1$Month
Xs.stats.JM$year <- data.jags.JM$data.1$Year
Xs.stats.JM$Season <- data.jags.JM$data.1$Season
Xs.stats.JM$location <- "Jamursba-Medi"

Xs.stats.W <- data.frame(low = as.vector(t(jm$q2.5$X[,,2])),
                         median = as.vector(t(jm$q50$X[,,2])),
                         high = as.vector(t(jm$q97.5$X[,,2])))
Xs.stats.W$location <- "Wermon"

Xs.stats.W$time <- data.jags.W$data.1$Frac.Year
Xs.stats.W$obsY <- data.jags.W$data.1$Nests
Xs.stats.W$month <- data.jags.W$data.1$Month
Xs.stats.W$year <- data.jags.W$data.1$Year
Xs.stats.W$Season <- data.jags.W$data.1$Season

Xs.stats <- rbind(Xs.stats.JM, Xs.stats.W)

Ns.stats.JM <- data.frame(time = year.begin:year.end,
                          low = as.vector(t(jm$q2.5$N[1,])),
                          median = as.vector(t(jm$q50$N[1,])),
                          high = as.vector(t(jm$q97.5$N[1,])))
Ns.stats.JM$location <- "Jamursba-Medi"

Ns.stats.W <- data.frame(time = year.begin:year.end,
                         low = as.vector(t(jm$q2.5$N[2,])),
                         median = as.vector(t(jm$q50$N[2,])),
                         high = as.vector(t(jm$q97.5$N[2,])))
Ns.stats.W$location <- "Wermon"

Ns.stats <- rbind(Ns.stats.JM, Ns.stats.W)
```


Save results if needed

```{r}
results.all <- list(jm = jm,
                    Xs.stats = Xs.stats,
                    ys.stats = ys.stats,
                    Ns.stats = Ns.stats)
if (save.data)
  saveRDS(results.all,
          file = paste0("RData/", filename.root, '.rds'))
```


Look at some posteriors
Look at some posteriors
```{r}
# pop growth rates - over the entire time
# bayesplot::mcmc_trace(jm$samples, c("U[1]", "U[2]"))
# 
# bayesplot::mcmc_trace(jm$samples, "p.beta.cos[1]")

bayesplot::mcmc_dens(jm$samples, c("U[1]", "U[2]"))

```

CV for two beaches:
```{r}
bayesplot::mcmc_dens(jm$samples, c("cv.Q[1]", "cv.Q[2]"))


```



```{r}
# cos and sin for the discrete Foureir series 
bayesplot::mcmc_dens(jm$samples, c("p.beta.cos[1]", "p.beta.sin[1]",
                                   "p.beta.cos[2]", "p.beta.sin[2]"))


```

```{r}
bayesplot::mcmc_dens(jm$samples, "sigma.N")

```

Posteror on log(N) for JM
```{r}
bayesplot::mcmc_dens(jm$samples, c("N[1,1]", "N[1,2]",
                                   "N[1,3]", "N[1,4]",
                                   "N[1,5]", "N[1,6]",
                                   "N[1,7]", "N[1,8]",
                                   "N[1,9]", "N[1,10]",
                                   "N[1,11]", "N[1,12]",
                                   "N[1,13]", "N[1,14]",
                                   "N[1,15]", "N[1,16]"))


```

Posteror on log(N) for W
```{r}
bayesplot::mcmc_dens(jm$samples, c("N[2,1]", "N[2,2]",
                                   "N[2,3]", "N[2,4]",
                                   "N[2,5]", "N[2,6]",
                                   "N[2,7]", "N[2,8]",
                                   "N[2,9]", "N[2,10]",
                                   "N[2,11]", "N[2,12]",
                                   "N[2,13]", "N[2,14]",
                                   "N[2,15]", "N[2,16]"))


```


Make some plots (need to change colors... )

```{r}

p.1a <- ggplot() +
  geom_ribbon(data = Xs.stats,
              aes(x = time, 
                  ymin = low, 
                  ymax = high),
              fill = fill.color,
              alpha = fill.alpha) + 
  geom_point(data = Xs.stats,
             aes(x = time, 
                 y = median),
             color = data.color,
             alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, 
                y = median),
            color = line.color,
            alpha = 0.5) +
  geom_point(data = ys.stats,
             aes(x = time, 
                 y = log(obsY)),
             color = obsd.color,
             alpha = 0.5) + 
  geom_ribbon(data = Ns.stats,
              aes(x = time, 
                  ymin = low,
                  ymax = high),
              color = fill.color.N1,
              alpha = fill.alpha) + 
  geom_line(data = Ns.stats,
            aes(x = time, y = median),
            color = line.color.N,
            alpha = 0.5,
            size = 1.5) + 
  scale_x_continuous(breaks = seq(year.begin, year.end, 5),
                     limits = c(year.begin, year.end)) +
  scale_y_continuous(limits = c(0, log(maxN))) + 
  facet_grid(rows = vars(location)) + 
  labs(x = '', y = 'log(# nests)')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

p.1a



```


```{r}
p.1b <- ggplot() +
  geom_ribbon(data = Xs.stats,
              aes(x = time, 
                  ymin = exp(low), 
                  ymax = exp(high)),
              fill = fill.color,
              alpha = fill.alpha) + 
  geom_point(data = Xs.stats,
             aes(x = time, 
                 y = exp(median)),
             color = data.color,
             alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, 
                y = exp(median)),
            color = line.color,
            alpha = 0.5) +
  geom_point(data = ys.stats,
             aes(x = time, 
                 y = obsY),
             color = obsd.color,
             alpha = 0.5) + 
  geom_ribbon(data = Ns.stats,
              aes(x = time+0.5, 
                  ymin = exp(low),
                  ymax = exp(high)),
              color = fill.color.N1,
              alpha = fill.alpha) + 
  geom_line(data = Ns.stats,
            aes(x = time+0.5, 
                y = exp(median)),
            color = line.color.N,
            alpha = 0.5,
            size = 1.5) + 
  geom_point(data = Ns.stats,
            aes(x = time+0.5, 
                y = exp(median)),
            color = "black",
            alpha = 0.5,
            size = 1.5) + 
  scale_x_continuous(breaks = seq(year.begin, year.end, 5),
                     limits = c(year.begin, year.end)) +
  scale_y_continuous(limits = c(0, 5000)) + 
  facet_grid(rows = vars(location)) + 
  labs(x = '', y = '# nests')  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

p.1b

```

save plots if requested

```{r}
if (save.fig){
  # ggsave(filename = paste0("figures/", filename.root, ".png"),
  #        plot = p.1,
  #        dpi = 600,
  #        device = "png")
  
  ggsave(filename = paste0("figures/", filename.root, ".png"),
         plot = p.1a,
         dpi = 600,
         device = "png")

  # ggsave(filename = paste0("figures/", filename.root,  "_pareto.png"),
  #      plot = p.2,
  #      dpi = 600,
  #      device = "png")
  
}

```

