
---
title: "Combining data imputation and trend analysis using discrete Fourier series"
output: html_notebook
---

Trying to merge the data imputation process and trend analysis with multiple datasets.

Model names were changed to streamline file names. Also added modified versions of models. In the new models (v2), the middle "state" space (x) is eliminated. Nt is modeled as a random variable, then Nt is divided into each month (X but not treated as a random variable), then y is observed from each X.

Set things up:

```{r}
rm(list=ls())

# tic <- Sys.time()
# Sys <- Sys.info()
source('Dc_Indonesia_nesting_fcns.R')
#library(rjags)

library(jagsUI)
library(coda)
library(tidyverse)
library(loo)
```


```{r}
save.data <- T
save.fig <- T
run.date <- "2020-10-14" #Sys.Date() #"2019-06-25" #

year.begin.JM <- 1999
year.begin.W <- 2006
#year.begin <- 2002
year.end <- 2019

period.JM <- 12
period.W <- 6
maxN <- 10000

UQ.models <- c("1U_1Q", 
               "1U_2Q",
               "2U_1Q",
               "2U_2Q")

# t_loc has location (or series) specific df parameter 
state.models <- c("norm", "t")
obs.models <- c("norm", "t") # 
v.models <- c("DFS_v1", "DFS_v2")  # v2 models have no Xs - Ns are the state

```


Set up the MCMC parameters

```{r}
MCMC.n.chains <- 5
MCMC.n.samples <- 50000
MCMC.n.burnin <- 10000
MCMC.n.thin <- 5

MCMC.params <- list(n.chains = MCMC.n.chains,
                    n.samples = MCMC.n.samples,
                    n.burnin = MCMC.n.burnin,
                    n.thin = MCMC.n.thin)
```


Bring in the data

```{r}
# Beginning year for JM is 1999
data.jags.JM <- data.extract(location = "JM", 
                             year.begin = year.begin.JM, 
                             year.end = year.end)

# Begining year for Wermon is 2006.
data.jags.W <- data.extract(location = "W", 
                             year.begin = year.begin.W, 
                             year.end = year.end)

```

Get warnings but they are harmless - I think. 

Combine datasets for analysis

```{r}
y <- array(data= NA, dim = c(nrow(data.jags.JM$jags.data2$y),
                             ncol(data.jags.JM$jags.data2$y), 2))
y[,,1] <- data.jags.JM$jags.data2$y
y[,,2] <- rbind(matrix(NA, nrow = nrow(data.jags.JM$jags.data2$y) - 
                         nrow(data.jags.W$jags.data2$y), ncol = 12),
                data.jags.W$jags.data2$y)

jags.data <- list(y = y,
                  #m = data.jags.JM$jags.data2$m,
                  C0 = c(15, 15),
                  n.months = 12,
                  C_cos = c(sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.JM))), 
                            sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.W)))),
                  C_sin = c(sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.JM))),
                            sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.W)))),
                  pi = pi,
                  period = c(period.JM, period.W),
                  N0_mean = c(8,8),   # this is 2980.958. Should be close enough
                  N0_sd = c(10, 10),
                  n.states = 2)

#n.timeseries <- dim(y)[3]

#jags.data$n.timeseries <- n.timeseries
jags.data$n.years <- dim(y)[1]

# jags.data$q_alpha <- 2
# jags.data$q_beta <- 0.5
# jags.data$r_alpha <- 2
# jags.data$r_beta <- 0.5

```

run jags on each model
```{r}
c <- 1
loo.out <- list()
filenames <- list()
k1 <- k2 <- 1
for (k1 in 1:length(UQ.models) ){ #3:()
  UQ.model <- UQ.models[k1]
  for (k2 in 1:length(obs.models)){
    obs.model <- obs.models[k2]
    for (k3 in 1:length(state.models)){
      state.model <- state.models[k3]
      for (k4 in 1:length(v.models)){
        v.model <- v.models[k4]
        filename.root <- paste0("SSAR1_", state.model, "_", 
                                obs.model, "_trend_", 
                                v.model, "_",
                                UQ.model, "_",
                                "JM", period.JM, 
                                "_W", period.W, "_",
                                year.begin.JM, "_", 
                                year.end, "_", run.date)
        
        filenames[[c]] <- filename.root
        # define parameters to monitor - add "df" for t distribution
        jags.params <- c('N', 'U', "p", "p.beta.cos", "p.beta.sin",
                         "sigma.Q", "sigma.R", 
                         "mu", "y", "deviance", "loglik")
        
        if (v.model == "DFS_v1"){
          jags.params <- c(jags.params, "X")
        }
        
        if (obs.model == "t"){
          jags.params <- c(jags.params, "df.y")
        }
        
        if (state.model == "t"){
          jags.params <- c(jags.params, "df.X")
        }
        
        model.name <- paste0("models/trend_models/model_", 
                             state.model, "_", obs.model, 
                             "_trend_", v.model, "_", 
                             UQ.model, ".txt")
        
        if (file.exists(model.name)){
          
          print(paste0("Running: ", model.name))
          
          if (!file.exists(paste0("RData/", filename.root, '.rds'))){
            jm <- jags(jags.data,
                       inits = NULL,
                       parameters.to.save= jags.params,
                       model.file = model.name,
                       n.chains = MCMC.params$n.chains,
                       n.burnin = MCMC.params$n.burnin,
                       n.thin = MCMC.params$n.thin,
                       n.iter = MCMC.params$n.samples,
                       DIC = T, parallel=T)
            
            saveRDS(jm,
                    file = paste0("RData/", filename.root, '.rds'))
          } else {
            jm <- readRDS(file = paste0("RData/", filename.root, '.rds'))
          }
          
          if (!file.exists(paste0("RData/", filename.root, "_loo.rds"))){
            loo.out[[c]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                          MCMC.params = MCMC.params, 
                                          data.vector = as.vector(jags.data$y))
            saveRDS(loo.out[[c]], file = paste0("RData/", filename.root, "_loo.rds"))
          } else {
            loo.out[[c]] <- readRDS(file = paste0("RData/", filename.root, "_loo.rds"))
            
          }
          
          c <- c + 1
          
        } else {
          print(paste0("Model file ", model.name, " does not exist" ))
        }    
        
      }
      
    }
    
  }
}

```

First look at the Pareto k diagnostics first to see how the models fit.

```{r}
pareto.k <- lapply(loo.out, 
                   FUN = function(x) x$loo.out)

# These have been saved so commenting out 2020-10-23
#saveRDS(loo.out, file = "RData/loo_all_Oct2020.rds")
#saveRDS(filenames, file = "RData/filenames_Oct2020.rds")

# find maximum pareto k values
max.pareto.k <- unlist(lapply(pareto.k,
       FUN = function(x) max(x$diagnostics$pareto_k)))

# find the models that have max(pareto k) < 0.7
good.models <- filenames[which(max.pareto.k < 0.7)]
good.models.pareto.k <- pareto.k[which(max.pareto.k < 0.7)]
```


Compare the results using LOOIC:

```{r}
looic.esimates <- lapply(lapply(loo.out[which(max.pareto.k < 0.7)], 
                                FUN = function(x) x$loo.out),
                         FUN = function(x) x$estimates)

looic <- unlist(lapply(looic.esimates, 
                       FUN = function(x) x["looic", "Estimate"]))

loo.out.list <- lapply(loo.out[which(max.pareto.k < 0.7)], 
                       FUN = function(x) x$loo.out)

# calculate model weights
model.weights <- loo_model_weights(loo.out.list)

model.names.abb <- lapply(good.models, 
                          FUN = function(x) strsplit(x, split = "SSAR1_")[[1]][2]) %>%  
  lapply(FUN = function(x) strsplit(x, split = "_JM12")[[1]][1]) %>% unlist() 

options(scipen = 999)

looic.table <- data.frame(model = model.names.abb,
                          looic = looic,
                          weights = as.vector(model.weights)) %>% 
  arrange(by = desc(weights)) %>%
  mutate(delta.looic = looic - min(looic)) %>%
  mutate_if(is.numeric, round, digits = 4)

best.model <- good.models[which(looic == min(looic))]
pareto.k.best <- good.models.pareto.k[[which(looic == min(looic))]]
```



```{r}
# two U model:
jm <- readRDS(file = paste0("RData/", best.model[[1]], ".rds"))
#max(unlist(lapply(jm$Rhat, FUN = max, na.rm = T)))
data.frame(jm$summary) %>% rownames_to_column("Parameter") -> summary.df

# check convergence 
rhat.max <- max(unlist(lapply(jm$Rhat, FUN = max, na.rm = T)))

# look at some posteriors
U.stats <- summary.df[grep(summary.df$Parameter, pattern = "U"),]
p.post.U <- bayesplot::mcmc_dens(jm$samples, c("U[1]", "U[2]")) + 
  xlab("Annual growth rate") + ylab("Density")

if (save.fig)
  ggsave(filename = 'figures/U_posterior_2Us.png', 
         plot = p.post.U,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```


```{r}
ys.stats.JM <- data.frame(low = as.vector(t(jm$q2.5$y[,,1])),
                          median = as.vector(t(jm$q50$y[,,1])),
                          high = as.vector(t(jm$q97.5$y[,,1])))

ys.stats.JM$time <- data.jags.JM$data.1$Frac.Year
ys.stats.JM$obsY <- data.jags.JM$data.1$Nests
ys.stats.JM$month <- data.jags.JM$data.1$Month
ys.stats.JM$year <- data.jags.JM$data.1$Year
ys.stats.JM$Season <- data.jags.JM$data.1$Season
ys.stats.JM$location <- "Jamursba-Medi"

# Wermon has shorter time series so obsY need NAs and other
# variables taken from JM datasets
ys.stats.W <- data.frame(low = as.vector(t(jm$q2.5$y[,,2])),
                         median = as.vector(t(jm$q50$y[,,2])),
                         high = as.vector(t(jm$q97.5$y[,,2])))
ys.stats.W$time <- data.jags.JM$data.1$Frac.Year
ys.stats.W$obsY <- c(rep(NA, nrow(data.jags.JM$data.1) - nrow(data.jags.W$data.1)), 
                     data.jags.W$data.1$Nests) #data.jags.W$data.1$Nests #
ys.stats.W$month <- data.jags.JM$data.1$Month
ys.stats.W$year <- data.jags.JM$data.1$Year
ys.stats.W$Season <- data.jags.JM$data.1$Season
ys.stats.W$location <- "Wermon"
          
ys.stats <- rbind(ys.stats.JM, ys.stats.W)

Ns.stats.JM <- data.frame(time = year.begin.JM:year.end,
                          low = as.vector(t(jm$q2.5$N[1,])),
                          median = as.vector(t(jm$q50$N[1,])),
                          high = as.vector(t(jm$q97.5$N[1,])))
Ns.stats.JM$location <- "Jamursba-Medi"
          
Ns.stats.W <- data.frame(time = year.begin.JM:year.end,
                         low = as.vector(t(jm$q2.5$N[2,])),
                         median = as.vector(t(jm$q50$N[2,])),
                         high = as.vector(t(jm$q97.5$N[2,])))
Ns.stats.W$location <- "Wermon"
          
Ns.stats <- rbind(Ns.stats.JM, Ns.stats.W)

p.estimated_y_N <- ggplot() + 
  geom_ribbon(data = ys.stats,
              aes(x = time, ymin = low, ymax = high),
              alpha = 0.4)+
  geom_path(data = ys.stats,
             aes(x = time, y = median)) + 
  geom_point(data = ys.stats,
             aes(x = time, y = median),
             color = "yellow") + 
  geom_point(data = ys.stats,
             aes(x = time, y = log(obsY)),
             color = "red") + 
  geom_ribbon(data = Ns.stats,
              aes(x = time+0.5, ymin = low, ymax = high),
              alpha = 0.4) +
  geom_path(data = Ns.stats,
            aes(x = time+0.5, y = median)) +
  geom_point(data = Ns.stats,
             aes(x = time+0.5, y = median),
             color = "yellow") +
  facet_grid(vars(location))

if (save.fig)
  ggsave(filename = 'figures/estimated_y_N_2Us.png', 
         plot = p.estimated_y_N,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```


# Season specific analyses

Enter NAs in Wermon winter data and run the models using period = 12 to see summer only analysis.

Do the same with winter only by NA in summer data... 

The imputation will be different... will that be a problem? 

I already have season-specific data. Using 6 months as the period, try running the same models to see what happens...

Summer only

```{r}
y <- array(data= NA, 
           dim = c(nrow(data.jags.JM$jags.data.summer$y),
                   ncol(data.jags.JM$jags.data.summer$y), 2))

y[,,1] <- data.jags.JM$jags.data.summer$y
y[,,2] <- rbind(matrix(NA, nrow = nrow(data.jags.JM$jags.data.summer$y) - 
                         nrow(data.jags.W$jags.data.summer$y), ncol = 6),
                data.jags.W$jags.data.summer$y)

period.JM <- period.W <- 6

jags.data.summer <- list(y = y,
                  #m = data.jags.JM$jags.data2$m,
                  C0 = c(15, 15),
                  n.months = 6,
                  C_cos = c(sum(apply(matrix(1:6, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.JM))), 
                            sum(apply(matrix(1:6, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.W)))),
                  C_sin = c(sum(apply(matrix(1:6, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.JM))),
                            sum(apply(matrix(1:6, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.W)))),
                  pi = pi,
                  period = c(period.JM, period.W),
                  N0_mean = c(8,8),   # this is 2980.958. Should be close enough
                  N0_sd = c(10, 10),
                  n.states = 2)

#n.timeseries <- dim(y)[3]

#jags.data$n.timeseries <- n.timeseries
jags.data.summer$n.years <- dim(y)[1]
```


run jags on each model
```{r}
c <- 1
loo.out.summer <- list()
filenames.summer <- list()
k1 <- k2 <- 1
for (k1 in 1:length(UQ.models) ){ #3:()
  UQ.model <- UQ.models[k1]
  for (k2 in 1:length(obs.models)){
    obs.model <- obs.models[k2]
    for (k3 in 1:length(state.models)){
      state.model <- state.models[k3]
      for (k4 in 1:length(v.models)){
        v.model <- v.models[k4]
        filename.root <- paste0("SSAR1_summer_", state.model, "_", 
                                obs.model, "_trend_", 
                                v.model, "_",
                                UQ.model, "_",
                                "JM", period.JM, 
                                "_W", period.W, "_",
                                year.begin.JM, "_", 
                                year.end, "_", run.date)
        
        filenames.summer[[c]] <- filename.root
        # define parameters to monitor - add "df" for t distribution
        jags.params <- c('N', 'U', "p", "p.beta.cos", "p.beta.sin",
                         "sigma.Q", "sigma.R", 
                         "mu", "y", "deviance", "loglik")
        
        if (v.model == "DFS_v1"){
          jags.params <- c(jags.params, "X")
        }
        
        if (obs.model == "t"){
          jags.params <- c(jags.params, "df.y")
        }
        
        if (state.model == "t"){
          jags.params <- c(jags.params, "df.X")
        }
        
        model.name <- paste0("models/trend_models/model_", 
                             state.model, "_", obs.model, 
                             "_trend_", v.model, "_", 
                             UQ.model, ".txt")
        
        if (file.exists(model.name)){
          
          print(paste0("Running: ", model.name))
          
          if (!file.exists(paste0("RData/", filename.root, '.rds'))){
            jm <- jags(jags.data.summer,
                       inits = NULL,
                       parameters.to.save= jags.params,
                       model.file = model.name,
                       n.chains = MCMC.params$n.chains,
                       n.burnin = MCMC.params$n.burnin,
                       n.thin = MCMC.params$n.thin,
                       n.iter = MCMC.params$n.samples,
                       DIC = T, parallel=T)
            
            saveRDS(jm,
                    file = paste0("RData/", filename.root, '.rds'))
          } else {
            jm <- readRDS(file = paste0("RData/", filename.root, '.rds'))
          }
          
          if (!file.exists(paste0("RData/", filename.root, "_loo.rds"))){
            loo.out.summer[[c]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                                 MCMC.params = MCMC.params, 
                                                 data.vector = as.vector(jags.data.summer$y))
            saveRDS(loo.out.summer[[c]], file = paste0("RData/", filename.root, "_loo.rds"))
          } else {
            loo.out.summer[[c]] <- readRDS(file = paste0("RData/", filename.root, "_loo.rds"))
            
          }
          
          c <- c + 1
          
        } else {
          print(paste0("Model file ", model.name, " does not exist" ))
        }    
        
      }
      
    }
    
  }
}

```

First look at the Pareto k diagnostics first to see how the models fit.

```{r}
pareto.k.summer <- lapply(loo.out.summer, 
                   FUN = function(x) x$loo.out)

# These have been saved so commenting out 2020-10-23
#saveRDS(loo.out.summer, file = "RData/loo_summer_Oct2020.rds")
#saveRDS(filenames.summer, file = "RData/filenames_summer_Oct2020.rds")

# find maximum pareto k values
max.pareto.k.summer <- unlist(lapply(pareto.k.summer,
       FUN = function(x) max(x$diagnostics$pareto_k)))

# find the models that have max(pareto k) < 0.7
good.models.summer <- filenames.summer[which(max.pareto.k.summer < 0.7)]
good.models.pareto.k.summer <- pareto.k.summer[which(max.pareto.k.summer < 0.7)]
```


Compare the results using LOOIC:

```{r}
looic.esimates.summer <- lapply(lapply(loo.out.summer[which(max.pareto.k.summer < 0.7)], 
                                FUN = function(x) x$loo.out),
                         FUN = function(x) x$estimates)

looic.summer <- unlist(lapply(looic.esimates.summer, 
                       FUN = function(x) x["looic", "Estimate"]))

loo.out.list.summer <- lapply(loo.out.summer[which(max.pareto.k.summer < 0.7)], 
                       FUN = function(x) x$loo.out)

# calculate model weights
model.weights.summer <- loo_model_weights(loo.out.list.summer)

model.names.abb.summer <- lapply(good.models.summer, 
                          FUN = function(x) strsplit(x, split = "SSAR1_")[[1]][2]) %>%  
  lapply(FUN = function(x) strsplit(x, split = "_JM6")[[1]][1]) %>% unlist() 

options(scipen = 999)

looic.table.summer <- data.frame(model = model.names.abb.summer,
                          looic = looic.summer,
                          weights = as.vector(model.weights.summer)) %>% 
  mutate(delta.looic = looic.summer - min(looic.summer)) %>%
  arrange(by = desc(weights)) %>%
  mutate_if(is.numeric, round, digits = 4)

best.model.summer <- good.models.summer[which(looic.summer == min(looic.summer))]
pareto.k.best.summer <- good.models.pareto.k.summer[[which(looic.summer == min(looic.summer))]]
```


Check posteriors
According to LOOIC, t_t_v2_1U_2Q is the best but t_t_v2_2U_2Q is best when using model weights. 
```{r}
# One U model:
jm <- readRDS(file = paste0("RData/", best.model.summer[[1]], ".rds"))
#max(unlist(lapply(jm$Rhat, FUN = max, na.rm = T)))
data.frame(jm$summary) %>% rownames_to_column("Parameter") -> summary.df

# check convergence 
rhat.max <- max(unlist(lapply(jm$Rhat, FUN = max, na.rm = T)))

# look at some posteriors
U.stats <- summary.df[grep(summary.df$Parameter, pattern = "U"),]
p.post.U <- bayesplot::mcmc_dens(jm$samples, c("U")) + 
  xlab("Annual growth rate") + ylab("Density")

if (save.fig)
  ggsave(filename = 'figures/U_posterior_summer_1U.png', 
         plot = p.post.U,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```


```{r}
ys.stats.JM <- data.frame(low = as.vector(t(jm$q2.5$y[,,1])),
                          median = as.vector(t(jm$q50$y[,,1])),
                          high = as.vector(t(jm$q97.5$y[,,1])))

ys.stats.JM$time <- data.jags.JM$data.summer$Frac.Year
ys.stats.JM$obsY <- data.jags.JM$data.summer$Nests
ys.stats.JM$month <- data.jags.JM$data.summer$Month
ys.stats.JM$year <- data.jags.JM$data.summer$Year
ys.stats.JM$Season <- data.jags.JM$data.summer$Season
ys.stats.JM$location <- "Jamursba-Medi"

# Wermon has shorter time series so obsY need NAs and other
# variables taken from JM datasets
ys.stats.W <- data.frame(low = as.vector(t(jm$q2.5$y[,,2])),
                         median = as.vector(t(jm$q50$y[,,2])),
                         high = as.vector(t(jm$q97.5$y[,,2])))
ys.stats.W$time <- data.jags.JM$data.summer$Frac.Year
ys.stats.W$obsY <- c(rep(NA, nrow(data.jags.JM$data.summer) - nrow(data.jags.W$data.summer)), 
                     data.jags.W$data.summer$Nests) #data.jags.W$data.1$Nests #
ys.stats.W$month <- data.jags.JM$data.summer$Month
ys.stats.W$year <- data.jags.JM$data.summer$Year
ys.stats.W$Season <- data.jags.JM$data.summer$Season
ys.stats.W$location <- "Wermon"
          
ys.stats <- rbind(ys.stats.JM, ys.stats.W)

Ns.stats.JM <- data.frame(time = year.begin.JM:year.end,
                          low = as.vector(t(jm$q2.5$N[1,])),
                          median = as.vector(t(jm$q50$N[1,])),
                          high = as.vector(t(jm$q97.5$N[1,])))
Ns.stats.JM$location <- "Jamursba-Medi"
          
Ns.stats.W <- data.frame(time = year.begin.JM:year.end,
                         low = as.vector(t(jm$q2.5$N[2,])),
                         median = as.vector(t(jm$q50$N[2,])),
                         high = as.vector(t(jm$q97.5$N[2,])))
Ns.stats.W$location <- "Wermon"
          
Ns.stats <- rbind(Ns.stats.JM, Ns.stats.W)

p.estimated_y_N <- ggplot() + 
  geom_ribbon(data = ys.stats,
              aes(x = time, ymin = low, ymax = high),
              alpha = 0.4)+
  geom_path(data = ys.stats,
             aes(x = time, y = median)) + 
  geom_point(data = ys.stats,
             aes(x = time, y = median),
             color = "yellow") + 
  geom_point(data = ys.stats,
             aes(x = time, y = log(obsY)),
             color = "red") + 
  geom_ribbon(data = Ns.stats,
              aes(x = time+0.5, ymin = low, ymax = high),
              alpha = 0.4) +
  geom_path(data = Ns.stats,
            aes(x = time+0.5, y = median)) +
  geom_point(data = Ns.stats,
             aes(x = time+0.5, y = median),
             color = "yellow") +
  facet_grid(vars(location))

if (save.fig)
  ggsave(filename = 'figures/estimated_y_N_summer_1U.png', 
         plot = p.estimated_y_N,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```


2 U model
```{r}
# Two U model:
jm <- readRDS(file = paste0("RData/SSAR1_", looic.table.summer[1,"model"], 
                            "_JM6_W6_1999_2019_2020-10-14.rds"))
#max(unlist(lapply(jm$Rhat, FUN = max, na.rm = T)))
data.frame(jm$summary) %>% rownames_to_column("Parameter") -> summary.df

# check convergence 
rhat.max <- max(unlist(lapply(jm$Rhat, FUN = max, na.rm = T)))

# look at some posteriors
U.stats <- summary.df[grep(summary.df$Parameter, pattern = "U"),]
p.post.U <- bayesplot::mcmc_dens(jm$samples, c("U[1]", "U[2]")) + 
  xlab("Annual growth rate") + ylab("Density")

if (save.fig)
  ggsave(filename = 'figures/U_posterior_summer_2Us.png', 
         plot = p.post.U,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```


```{r}
ys.stats.JM <- data.frame(low = as.vector(t(jm$q2.5$y[,,1])),
                          median = as.vector(t(jm$q50$y[,,1])),
                          high = as.vector(t(jm$q97.5$y[,,1])))

ys.stats.JM$time <- data.jags.JM$data.summer$Frac.Year
ys.stats.JM$obsY <- data.jags.JM$data.summer$Nests
ys.stats.JM$month <- data.jags.JM$data.summer$Month
ys.stats.JM$year <- data.jags.JM$data.summer$Year
ys.stats.JM$Season <- data.jags.JM$data.summer$Season
ys.stats.JM$location <- "Jamursba-Medi"

# Wermon has shorter time series so obsY need NAs and other
# variables taken from JM datasets
ys.stats.W <- data.frame(low = as.vector(t(jm$q2.5$y[,,2])),
                         median = as.vector(t(jm$q50$y[,,2])),
                         high = as.vector(t(jm$q97.5$y[,,2])))
ys.stats.W$time <- data.jags.JM$data.summer$Frac.Year
ys.stats.W$obsY <- c(rep(NA, nrow(data.jags.JM$data.summer) - nrow(data.jags.W$data.summer)), 
                     data.jags.W$data.summer$Nests) #data.jags.W$data.1$Nests #
ys.stats.W$month <- data.jags.JM$data.summer$Month
ys.stats.W$year <- data.jags.JM$data.summer$Year
ys.stats.W$Season <- data.jags.JM$data.summer$Season
ys.stats.W$location <- "Wermon"
          
ys.stats <- rbind(ys.stats.JM, ys.stats.W)

Ns.stats.JM <- data.frame(time = year.begin.JM:year.end,
                          low = as.vector(t(jm$q2.5$N[1,])),
                          median = as.vector(t(jm$q50$N[1,])),
                          high = as.vector(t(jm$q97.5$N[1,])))
Ns.stats.JM$location <- "Jamursba-Medi"
          
Ns.stats.W <- data.frame(time = year.begin.JM:year.end,
                         low = as.vector(t(jm$q2.5$N[2,])),
                         median = as.vector(t(jm$q50$N[2,])),
                         high = as.vector(t(jm$q97.5$N[2,])))
Ns.stats.W$location <- "Wermon"
          
Ns.stats <- rbind(Ns.stats.JM, Ns.stats.W)

p.estimated_y_N <- ggplot() + 
  geom_ribbon(data = ys.stats,
              aes(x = time, ymin = low, ymax = high),
              alpha = 0.4)+
  geom_path(data = ys.stats,
             aes(x = time, y = median)) + 
  geom_point(data = ys.stats,
             aes(x = time, y = median),
             color = "yellow") + 
  geom_point(data = ys.stats,
             aes(x = time, y = log(obsY)),
             color = "red") + 
  geom_ribbon(data = Ns.stats,
              aes(x = time+0.5, ymin = low, ymax = high),
              alpha = 0.4) +
  geom_path(data = Ns.stats,
            aes(x = time+0.5, y = median)) +
  geom_point(data = Ns.stats,
             aes(x = time+0.5, y = median),
             color = "yellow") +
  facet_grid(vars(location))

if (save.fig)
  ggsave(filename = 'figures/estimated_y_N_summer_2Us.png', 
         plot = p.estimated_y_N,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```

Wermon winter

Only one U models are needed.

```{r}
UQ.models <- c("1U_1Q", 
               "1U_2Q")
```


```{r}
y <- array(data= NA, dim = c(nrow(data.jags.W$jags.data.winter$y),
                             ncol(data.jags.W$jags.data.winter$y), 1))
y[,,1] <- data.jags.W$jags.data.winter$y

period.W <- 6

jags.data.winter <- list(y = y,
                         #m = data.jags.JM$jags.data2$m,
                         C0 = c(15, 15),
                         n.months = 6,
                         C_cos = sum(apply(matrix(1:6, nrow=1), 
                                           MARGIN = 1, 
                                           FUN = function(x) cos(2 * pi * x/period.W))),
                         C_sin = sum(apply(matrix(1:6, nrow=1), 
                                           MARGIN = 1, 
                                           FUN = function(x) sin(2 * pi * x/period.W))),
                         pi = pi,
                         period = 6,
                         N0_mean = c(8,8),   # this is 2980.958. Should be close enough
                         N0_sd = c(10, 10),
                         n.states = 1)

#n.timeseries <- dim(y)[3]

#jags.data$n.timeseries <- n.timeseries
jags.data.winter$n.years <- dim(y)[1]
```


run jags on each model
```{r}
c <- 1
loo.out.winter <- list()
filenames.winter <- list()
k1 <- k2 <- k3 <- k4 <- 1
for (k1 in 1:length(UQ.models) ){ #3:()
  UQ.model <- UQ.models[k1]
  for (k2 in 1:length(obs.models)){
    obs.model <- obs.models[k2]
    for (k3 in 1:length(state.models)){
      state.model <- state.models[k3]
      for (k4 in 1:length(v.models)){
        v.model <- v.models[k4]
        filename.root <- paste0("SSAR1_winter_", state.model, "_", 
                                obs.model, "_trend_", 
                                v.model, "_",
                                UQ.model, "_",
                                "W", period.W, "_",
                                year.begin.W, "_", 
                                year.end, "_", run.date)
        
        filenames.winter[[c]] <- filename.root
        # define parameters to monitor - add "df" for t distribution
        jags.params <- c('N', 'U', "p", "p.beta.cos", "p.beta.sin",
                         "sigma.Q", "sigma.R", 
                         "mu", "y", "deviance", "loglik")
        
        if (v.model == "DFS_v1"){
          jags.params <- c(jags.params, "X")
        }
        
        if (obs.model == "t"){
          jags.params <- c(jags.params, "df.y")
        }
        
        if (state.model == "t"){
          jags.params <- c(jags.params, "df.X")
        }
        
        model.name <- paste0("models/trend_models/model_", 
                             state.model, "_", obs.model, 
                             "_trend_", v.model, "_", 
                             UQ.model, ".txt")
        
        if (file.exists(model.name)){
          
          print(paste0("Running: ", model.name))
          
          if (!file.exists(paste0("RData/", filename.root, '.rds'))){
            jm <- jags(data = jags.data.winter,
                       inits = NULL,
                       parameters.to.save= jags.params,
                       model.file = model.name,
                       n.chains = MCMC.params$n.chains,
                       n.burnin = MCMC.params$n.burnin,
                       n.thin = MCMC.params$n.thin,
                       n.iter = MCMC.params$n.samples,
                       DIC = T, parallel=T)
            
            saveRDS(jm,
                    file = paste0("RData/", filename.root, '.rds'))
          } else {
            jm <- readRDS(file = paste0("RData/", filename.root, '.rds'))
          }
          
          if (!file.exists(paste0("RData/", filename.root, "_loo.rds"))){
            loo.out.winter[[c]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                                 MCMC.params = MCMC.params, 
                                                 data.vector = as.vector(jags.data.winter$y))
            saveRDS(loo.out.winter[[c]], file = paste0("RData/", filename.root, "_loo.rds"))
          } else {
            loo.out.winter[[c]] <- readRDS(file = paste0("RData/", filename.root, "_loo.rds"))
            
          }
          
          c <- c + 1
          
        } else {
          print(paste0("Model file ", model.name, " does not exist" ))
        }    
        
      }
      
    }
    
  }
}

```

First look at the Pareto k diagnostics first to see how the models fit.

```{r}
pareto.k.winter <- lapply(loo.out.winter, 
                   FUN = function(x) x$loo.out)

# These have been saved so commenting out 2020-10-23
#saveRDS(loo.out.winter, file = "RData/loo_winter_Oct2020.rds")
#saveRDS(filenames.winter, file = "RData/filenames_winter_Oct2020.rds")

# find maximum pareto k values
max.pareto.k.winter <- unlist(lapply(pareto.k.winter,
       FUN = function(x) max(x$diagnostics$pareto_k)))

# find the models that have max(pareto k) < 0.7
good.models.winter <- filenames.winter[which(max.pareto.k.winter < 0.7)]
good.models.pareto.k.winter <- pareto.k.winter[which(max.pareto.k.winter < 0.7)]

best.model.winter <- good.models.winter
pareto.k.best.winter <- good.models.pareto.k.winter

```

Only one model fitted okay... So, no need to compare models.

<!-- Compare the results using LOOIC: -->

<!-- ```{r} -->
<!-- looic.esimates.summer <- lapply(lapply(loo.out.summer[which(max.pareto.k.summer < 0.7)],  -->
<!--                                 FUN = function(x) x$loo.out), -->
<!--                          FUN = function(x) x$estimates) -->

<!-- looic.summer <- unlist(lapply(looic.esimates.summer,  -->
<!--                        FUN = function(x) x["looic", "Estimate"])) -->

<!-- loo.out.list.summer <- lapply(loo.out.summer[which(max.pareto.k.summer < 0.7)],  -->
<!--                        FUN = function(x) x$loo.out) -->

<!-- # calculate model weights -->
<!-- model.weights.summer <- loo_model_weights(loo.out.list.summer) -->

<!-- model.names.abb.summer <- lapply(good.models.summer,  -->
<!--                           FUN = function(x) strsplit(x, split = "SSAR1_")[[1]][2]) %>%   -->
<!--   lapply(FUN = function(x) strsplit(x, split = "_JM6")[[1]][1]) %>% unlist()  -->

<!-- options(scipen = 999) -->

<!-- looic.table.summer <- data.frame(model = model.names.abb.summer, -->
<!--                           looic = looic.summer, -->
<!--                           weights = as.vector(model.weights.summer)) %>%  -->
<!--   arrange(by = desc(weights)) %>% -->
<!--   mutate(delta.looic = looic.summer - min(looic.summer)) %>% -->
<!--   mutate_if(is.numeric, round, digits = 4) -->

<!-- best.model.summer <- good.models.summer[which(looic.summer == min(looic.summer))] -->
<!-- pareto.k.best.summer <- good.models.pareto.k.summer[[which(looic.summer == min(looic.summer))]] -->
<!-- ``` -->


Check posteriors

```{r}
jm <- readRDS(file = paste0("RData/", best.model.winter[[1]], ".rds"))
data.frame(jm$summary) %>% rownames_to_column("Parameter") -> summary.df

# check convergence 
rhat.max <- max(unlist(lapply(jm$Rhat, FUN = max, na.rm = T)))

# look at some posteriors
U.stats <- summary.df[grep(summary.df$Parameter, pattern = "U"),]
p.post.U <- bayesplot::mcmc_dens(jm$samples, c("U")) + 
  xlab("Annual growth rate") + ylab("Density")

if (save.fig)
  ggsave(filename = 'figures/U_posterior_winter.png', 
         plot = p.post.U,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```


```{r}

ys.stats <- data.frame(low = as.vector(t(jm$q2.5$y[,,1])),
                         median = as.vector(t(jm$q50$y[,,1])),
                         high = as.vector(t(jm$q97.5$y[,,1])))
ys.stats$time <- data.jags.W$data.winter$Frac.Year
ys.stats$obsY <- data.jags.W$data.winter$Nests
ys.stats$month <- data.jags.W$data.winter$Month
ys.stats$year <- data.jags.W$data.winter$Year
ys.stats$Season <- data.jags.W$data.winter$Season
ys.stats$location <- "Wermon"
          
Ns.stats <- data.frame(time = year.begin.W:year.end,
                         low = as.vector(t(jm$q2.5$N[1,])),
                         median = as.vector(t(jm$q50$N[1,])),
                         high = as.vector(t(jm$q97.5$N[1,])))
Ns.stats$location <- "Wermon"
          
p.estimated_y_N <- ggplot() + 
  geom_ribbon(data = ys.stats,
              aes(x = time, ymin = low, ymax = high),
              alpha = 0.4)+
  geom_path(data = ys.stats,
             aes(x = time, y = median)) + 
  geom_point(data = ys.stats,
             aes(x = time, y = median),
             color = "yellow") + 
  geom_point(data = ys.stats,
             aes(x = time, y = log(obsY)),
             color = "red") + 
  geom_ribbon(data = Ns.stats,
              aes(x = time+0.5, ymin = low, ymax = high),
              alpha = 0.4) +
  geom_path(data = Ns.stats,
            aes(x = time+0.5, y = median)) +
  geom_point(data = Ns.stats,
             aes(x = time+0.5, y = median),
             color = "yellow") 

if (save.fig)
  ggsave(filename = 'figures/estimated_y_N_winter.png', 
         plot = p.estimated_y_N,
         dpi = 600, height = 6, 
         width = 8, units = "in")
```

