
---
title: "Combining data imputation and trend analysis using discrete Fourier series"
output: html_notebook
---

Trying to merge the data imputation process and trend analysis with multiple datasets.

Set things up:

```{r}
rm(list=ls())

# tic <- Sys.time()
# Sys <- Sys.info()
source('Dc_Indonesia_nesting_fcns.R')
#library(rjags)

library(jagsUI)
library(coda)
library(tidyverse)
library(loo)
```


```{r}
save.data <- T
save.fig <- T
run.date <- "2020-09-24" #Sys.Date() #"2019-06-25" #

year.begin <- 2001
year.end <- 2019

period.JM <- 12
period.W <- 6
maxN <- 10000

model.names <- c("singleUQ", 
                 "singleU_independentQs",
                 "independentUQs",
                 "independentUs_singleQ")

# t_loc has location (or series) specific df parameter 
obs.models <- c("norm", "t") # 

```


Set up the MCMC parameters

```{r}
MCMC.n.chains <- 5
MCMC.n.samples <- 50000
MCMC.n.burnin <- 10000
MCMC.n.thin <- 5

MCMC.params <- list(n.chains = MCMC.n.chains,
                    n.samples = MCMC.n.samples,
                    n.burnin = MCMC.n.burnin,
                    n.thin = MCMC.n.thin)
```


Bring in the data

```{r}

data.jags.JM <- data.extract(location = "JM", 
                             year.begin = year.begin, 
                             year.end = year.end)

data.jags.W <- data.extract(location = "W", 
                             year.begin = year.begin, 
                             year.end = year.end)

```

Get warnings but they are harmless - I think. 

Combine datasets for analysis

```{r}
y <- array(data= NA, dim = c(nrow(data.jags.JM$jags.data2$y),
                             ncol(data.jags.JM$jags.data2$y), 2))
y[,,1] <- data.jags.JM$jags.data2$y
y[,,2] <- data.jags.W$jags.data2$y

jags.data <- list(y = y,
                  #m = data.jags.JM$jags.data2$m,
                  C0 = c(15, 15),
                  n.months = 12,
                  C_cos = c(sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.JM))), 
                            sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.W)))),
                  C_sin = c(sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.JM))),
                            sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.W)))),
                  pi = pi,
                  period = c(period.JM, period.W),
                  N0_mean = c(log(sum(exp(y[1,,1]), na.rm = T)),
                              log(sum(exp(y[1,,2]), na.rm = T))),
                  N0_sd = c(10, 10))

n.timeseries <- dim(y)[3]

jags.data$n.timeseries <- n.timeseries
jags.data$n.years <- dim(y)[1]

# jags.data$q_alpha <- 2
# jags.data$q_beta <- 0.5
# jags.data$r_alpha <- 2
# jags.data$r_beta <- 0.5

```

run jags on each model
```{r}
c <- 1
loo.out <- list()
filenames <- list()
k1 <- k2 <- 1
for (k1 in 1:length(model.names) ){ #3:()
  model.number <- k1
  model.name <- model.names[k1]

  # Model-specific parameters
  
  if (length(grep("independentU", model.name)) == 1){
    # multiple U
    whichPop <- 1:n.timeseries         # multiple time series -> unique population processes
    n.states <- max(whichPop)
    # 
    # # matrix with rows as n.timeseries and cols as n.states (pops)
    # Z <- matrix(0,n.timeseries+1,n.states+1)   
    # # add a row of NAs to keep jagsUI from converting single time series matrix into vector
    # Z[n.timeseries+1, ] <- NA                  
    # # add a col of NAs to keep jagsUI from converting single state matrix into vector
    # Z[ , n.states+1] <- NA                     
    # for(i in 1:length(whichPop)) Z[i,whichPop[i]] <- 1
    
  } else if (length(grep("singleU", model.name)) == 1){
    ### single U
    n.states <- 1
    # whichPop <- rep(1, n.timeseries)
    # # matrix with rows as n.timeseries and cols as n.states (pops)
    # Z <- matrix(0, n.timeseries+1, n.states+1) 
    # # add a row of NAs to keep jagsUI from converting single time series matrix into vector
    # Z[n.timeseries+1, ] <- NA                  
    # # add a col of NAs to keep jagsUI from converting single state matrix into vector
    # Z[ , n.states+1] <- NA                     
    # for(i in 1:n.timeseries) Z[i, whichPop[i]] <- 1
    
  }
  
#  jags.data$Z <- Z
  jags.data$n.states <- n.states

  for (k2 in 1:length(obs.models)){
    
    obs.model <- obs.models[k2]
    filename.root <- paste0("SSAR1_norm_", obs.models[k2], "_trend_Four_", 
                            model.names[model.number], "_",
                            "JM", period.JM, "_W", period.W, "_",
                            year.begin, "_", year.end, "_", run.date)
    
    filenames[[c]] <- filename.root
    # define parameters to monitor - add "df" for t distribution
    jags.params <- c('N', 'U', "p", "p.beta.cos", "p.beta.sin",
                     "sigma.N", 'sigma.Q', "sigma.R", 
                     "mu", "y", "X", "deviance", "loglik")
    
    if (obs.model == "t"){
      jags.params <- c(jags.params, "df")
    }
    
    if (length(grep("FourQ", model.names[model.number])) != 0){
      jags.params <- c(jags.params, c("C0_Q", "q.beta.cos", "q.beta.sin"))
    }
   
    if (length(grep("constCV", model.names[model.number])) != 0)
      jags.params <- c(jags.params, "cv.Q")
    
    model.name <- paste0("models/model_norm_", obs.model, "_trend_Four_", 
                     model.names[model.number], ".txt")
    
    if (file.exists(model.name)){
      
      if (!file.exists(paste0("RData/", filename.root, '.rds'))){
        jm <- jags(jags.data,
                 inits = NULL,
                 parameters.to.save= jags.params,
                 model.file = model.name,
                 n.chains = MCMC.params$n.chains,
                 n.burnin = MCMC.params$n.burnin,
                 n.thin = MCMC.params$n.thin,
                 n.iter = MCMC.params$n.samples,
                 DIC = T, parallel=T)
      
        ys.stats.JM <- data.frame(low = as.vector(t(jm$q2.5$y[,,1])),
                                  median = as.vector(t(jm$q50$y[,,1])),
                                  high = as.vector(t(jm$q97.5$y[,,1])))
        ys.stats.JM$time <- data.jags.JM$data.1$Frac.Year
        ys.stats.JM$obsY <- data.jags.JM$data.1$Nests
        ys.stats.JM$month <- data.jags.JM$data.1$Month
        ys.stats.JM$year <- data.jags.JM$data.1$Year
        ys.stats.JM$Season <- data.jags.JM$data.1$Season
        ys.stats.JM$location <- "Jamursba-Medi"
        
        ys.stats.W <- data.frame(low = as.vector(t(jm$q2.5$y[,,2])),
                                 median = as.vector(t(jm$q50$y[,,2])),
                                 high = as.vector(t(jm$q97.5$y[,,2])))
        ys.stats.W$time <- data.jags.JM$data.1$Frac.Year
        ys.stats.W$obsY <- c(rep(NA, 9), data.jags.W$data.1$Nests)
        ys.stats.W$month <- data.jags.JM$data.1$Month
        ys.stats.W$year <- data.jags.JM$data.1$Year
        ys.stats.W$Season <- data.jags.JM$data.1$Season
        ys.stats.W$location <- "Wermon"
        
        ys.stats <- rbind(ys.stats.JM, ys.stats.W)
        
        # extract Xs - the state model
        
        Xs.stats.JM <- data.frame(low = as.vector(t(jm$q2.5$X[,,1])),
                                  median = as.vector(t(jm$q50$X[,,1])),
                                  high = as.vector(t(jm$q97.5$X[,,1])))
        Xs.stats.JM$time <- data.jags.JM$data.1$Frac.Year
        Xs.stats.JM$obsY <- data.jags.JM$data.1$Nests
        Xs.stats.JM$month <- data.jags.JM$data.1$Month
        Xs.stats.JM$year <- data.jags.JM$data.1$Year
        Xs.stats.JM$Season <- data.jags.JM$data.1$Season
        Xs.stats.JM$location <- "Jamursba-Medi"
        
        Xs.stats.W <- data.frame(low = as.vector(t(jm$q2.5$X[,,2])),
                                 median = as.vector(t(jm$q50$X[,,2])),
                                 high = as.vector(t(jm$q97.5$X[,,2])))
        Xs.stats.W$location <- "Wermon"
        
        Xs.stats.W$time <- data.jags.JM$data.1$Frac.Year
        Xs.stats.W$obsY <- c(rep(NA, 9), data.jags.W$data.1$Nests)
        Xs.stats.W$month <- data.jags.JM$data.1$Month
        Xs.stats.W$year <- data.jags.JM$data.1$Year
        Xs.stats.W$Season <- data.jags.JM$data.1$Season
        
        Xs.stats <- rbind(Xs.stats.JM, Xs.stats.W)
        
        Ns.stats.JM <- data.frame(time = year.begin:year.end,
                                  low = as.vector(t(jm$q2.5$N[1,])),
                                  median = as.vector(t(jm$q50$N[1,])),
                                  high = as.vector(t(jm$q97.5$N[1,])))
        Ns.stats.JM$location <- "Jamursba-Medi"
        
        Ns.stats.W <- data.frame(time = year.begin:year.end,
                                 low = as.vector(t(jm$q2.5$N[2,])),
                                 median = as.vector(t(jm$q50$N[2,])),
                                 high = as.vector(t(jm$q97.5$N[2,])))
        Ns.stats.W$location <- "Wermon"
        
        Ns.stats <- rbind(Ns.stats.JM, Ns.stats.W)
        
        results.all <- list(jm = jm,
                            Xs.stats = Xs.stats,
                            ys.stats = ys.stats,
                            Ns.stats = Ns.stats)
        
        saveRDS(results.all,
                file = paste0("RData/", filename.root, '.rds'))
      } else {
        results.all <- readRDS(file = paste0("RData/", filename.root, '.rds'))
      }
      
      if (!file.exists(paste0("RData/", filename.root, "_loo.rds"))){
        jm <- results.all$jm
        
        loo.out[[c]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                      MCMC.params = MCMC.params, 
                                      data.vector = as.vector(jags.data$y))
        saveRDS(loo.out[[c]], file = paste0("RData/", filename.root, "_loo.rds"))
      } else {
        loo.out[[c]] <- readRDS(file = paste0("RData/", filename.root, "_loo.rds"))
        
      }
      
      c <- c + 1
      
    } else {
      print(paste0("Model file", model.name, " does not exist" ))
    }    

    
  }
  
}


```

Compare the results using LOOIC:

```{r}

looic.esimates <- lapply(lapply(loo.out, 
                                FUN = function(x) x$loo.out),
                         FUN = function(x) x$estimates)

looic <- unlist(lapply(looic.esimates, 
                       FUN = function(x) x["looic", "Estimate"]))


best.model <- filenames[which(looic == min(looic))]
```


According to the LOOIC values, normal process and normal observation models with independent slope and variance terms was considered best. Pareto k values, however, were not stellar... 

```{r}
pareto.k <- lapply(loo.out, 
                   FUN = function(x) x$loo.out)
pareto.k.best <- pareto.k[[which(looic == min(looic))]]
```

Take a look at which data points are problematic

```{r}

data.jags.JM$data.1 %>% select(Frac.Year, Nests, Season) %>%
  mutate(Loc = "JM") %>%
  # transmute(Frac.Year = Frac.Year,
  #           Nests.JM = Nests,
  #           Season = Season) %>% 
  na.omit() -> data.JM

data.jags.W$data.1 %>% select(Frac.Year, Nests, Season) %>%
  mutate(Loc = "W") %>%
  # transmute(Frac.Year = Frac.Year, 
  #           Nests.W = Nests) %>%
  na.omit() -> data.W

data.loo <- rbind(data.JM, data.W)
# n.JM <- nrow(data.JM)
# n.both <- n.JM + nrow(data.W)
data.loo$pareto.k <- pareto.k.best$diagnostics$pareto_k

ggplot(data = data.loo) + 
  geom_point(aes(x = Frac.Year, y = pareto.k, color = Loc))
```

Not great... but should we just move forward rather than trying to see I can come up with a better model? 

```{r}

jags.out <- readRDS(file = paste0("RData/", best.model[[1]], ".rds"))
jm <- jags.out$jm

bayesplot::mcmc_dens(jm$samples, c("U[1]", "U[2]"))
```

```{r}
data.frame(jm$summary) %>% rownames_to_column("Parameter") -> summary.df

summary.df[grep(summary.df$Parameter, pattern = "U\\["),]
```

These are almost identical... I'm not sure why the independent U model was considered best. 

```{r}
bayesplot::mcmc_dens(jm$samples, c("sigma.Q[1]", "sigma.Q[2]"))
```

Looks a bit suspicious...

```{r}
bayesplot::mcmc_trace(jm$samples, c("sigma.Q[1]", "sigma.Q[2]"))
```

They look alright. 

Let's look at single U model to see how the estimate is different - should be about the same.

```{r}
jags.out.1 <- readRDS(file = paste0("RData/", filenames[[3]], ".rds")) #"RData/SSAR1_norm_norm_trend_Four_singleU_independentQs_JM12_W6_2001-2019_2020-09-24.rds")

jm.1 <- jags.out.1$jm
data.frame(jm.1$summary) %>% rownames_to_column("Parameter") -> summary.1.df

grep(summary.df$Parameter, pattern = "U", fixed = T)
summary.1.df %>% filter(Parameter == "U")

```

```{r}
bayesplot::mcmc_dens(jm.1$samples, "U")
```

So, this is pretty much identical to the previous model (independent Us). LOOIC value for this one was only 2 greater than the other model (985.89 vs. 988.08). So, it may make sense to use the simpler model.

Pareto K?

```{r}
pareto.k[[3]]
```

It's about the same... 

```{r}

```

