
---
title: "Combining data imputation and trend analysis using discrete Fourier series for JM"
output: html_notebook
---

Trying to merge the data imputation process and trend analysis with one dataset.

Set things up:

```{r}
rm(list=ls())

# tic <- Sys.time()
# Sys <- Sys.info()
source('Dc_Indonesia_nesting_fcns.R')
#library(rjags)

library(jagsUI)
library(coda)
library(tidyverse)
library(loo)
```


```{r}
save.data <- T
save.fig <- T
run.date <- "2020-10-09" #Sys.Date() #"2019-06-25" #

# for JM, we can start early
year.begin <- 2002
year.end <- 2019

period.JM <- 12
period.W <- 6
maxN <- 10000

# t_loc has location (or series) specific df parameter 
state.models <- c("norm", "t")
obs.models <- c("norm", "t") # 
v2.models <- c("Four", "Four_v2")  # v2 models have no Xs - Ns are the state

```


Set up the MCMC parameters

```{r}
MCMC.n.chains <- 5
MCMC.n.samples <- 50000
MCMC.n.burnin <- 10000
MCMC.n.thin <- 5

MCMC.params <- list(n.chains = MCMC.n.chains,
                    n.samples = MCMC.n.samples,
                    n.burnin = MCMC.n.burnin,
                    n.thin = MCMC.n.thin)
```


Bring in the data

```{r}

data.jags.W <- data.extract(location = "W", 
                             year.begin = year.begin, 
                             year.end = year.end)

```

Combine datasets for analysis

```{r}
jags.data <- list(y = data.jags.W$jags.data2$y,
                  #m = data.jags.W$jags.data2$m,
                  C0 = 15,
                  n.months = 12,
                  C_cos = sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) cos(2 * pi * x/period.JM))),
                  C_sin = sum(apply(matrix(1:12, nrow=1), 
                                      MARGIN = 1, 
                                      FUN = function(x) sin(2 * pi * x/period.JM))),
                  pi = pi,
                  period = period.JM,
                  N0_mean = log(sum(exp(data.jags.W$jags.data2$y[1,]), na.rm = T)),
                  N0_sd = 10,
                  n.years = nrow(data.jags.W$jags.data2$y))

```

run jags on each model
```{r}
c <- 1
loo.out <- list()
filenames <- list()
k1 <- k2 <- k3 <- k4 <- 1

for (k2 in 1:length(obs.models)){
  for (k3 in 1:length(state.models)){
    for (k4 in 1:length(v2.models)){
      obs.model <- obs.models[k2]
      state.model <- state.models[k3]
          v2.model <- v2.models[k4]
      filename.root <- paste0("SSAR1_", state.model, "_", obs.model, "_trend_", v2.model, "_W", 
                            period.W, "_",
                            year.begin, "_", year.end, "_", run.date)
    
      filenames[[c]] <- filename.root
      # define parameters to monitor - add "df" for t distribution
      jags.params <- c("N", "theta", "p", "p.beta.cos", "p.beta.sin",
                       "sigma.N", 'sigma.pro', "sigma.obs", 
                       "mu", "y", "X", "deviance", "loglik")
      
      if (obs.model == "t"){
        jags.params <- c(jags.params, "df_y")
      }
      
      if (state.model == "t"){
        jags.params <- c(jags.params, "df_X")
        if (v2.model == "Four_v2"){
          jags.params <- c(jags.params, "df_N")
        }  
      }
      
      model.name <- paste0("models/model_", state.model, "_", 
                           obs.model, "_trend_", v2.model, ".txt")
      
      if (file.exists(model.name)){
        
        if (!file.exists(paste0("RData/", filename.root, '.rds'))){
          jm <- jags(jags.data,
                     inits = NULL,
                     parameters.to.save= jags.params,
                     model.file = model.name,
                     n.chains = MCMC.params$n.chains,
                     n.burnin = MCMC.params$n.burnin,
                     n.thin = MCMC.params$n.thin,
                     n.iter = MCMC.params$n.samples,
                     DIC = T, parallel=T)
          
          ys.stats <- data.frame(low = as.vector(t(jm$q2.5$y)),
                                 median = as.vector(t(jm$q50$y)),
                                 high = as.vector(t(jm$q97.5$y)))
          ys.stats$time <- data.jags.W$data.1$Frac.Year
          ys.stats$obsY <- data.jags.W$data.1$Nests
          ys.stats$month <- data.jags.W$data.1$Month
          ys.stats$year <- data.jags.W$data.1$Year
          ys.stats$Season <- data.jags.W$data.1$Season
          ys.stats$location <- "Wermon"
        
          if (v2.model == "Four"){
            
            # extract Xs - the state model
            
            Xs.stats <- data.frame(low = as.vector(t(jm$q2.5$X)),
                                   median = as.vector(t(jm$q50$X)),
                                   high = as.vector(t(jm$q97.5$X)))
            Xs.stats$time <- data.jags.W$data.1$Frac.Year
            Xs.stats$obsY <- data.jags.W$data.1$Nests
            Xs.stats$month <- data.jags.W$data.1$Month
            Xs.stats$year <- data.jags.W$data.1$Year
            Xs.stats$Season <- data.jags.W$data.1$Season
            Xs.stats$location <- "Wermon"
          }      
          Ns.stats <- data.frame(time = year.begin:year.end,
                                 low = as.vector(t(jm$q2.5$N)),
                                 median = as.vector(t(jm$q50$N)),
                                 high = as.vector(t(jm$q97.5$N)))
          Ns.stats$location <- "Wermon"
          
          if (v2.model == "Four"){
            results.all <- list(jm = jm,
                                Xs.stats = Xs.stats,
                                ys.stats = ys.stats,
                                Ns.stats = Ns.stats)
            
          } else {
            results.all <- list(jm = jm,
                                ys.stats = ys.stats,
                                Ns.stats = Ns.stats)
            
          }
          
          saveRDS(results.all,
                  file = paste0("RData/", filename.root, '.rds'))
        } else {
          results.all <- readRDS(file = paste0("RData/", filename.root, '.rds'))
        }
        
        if (!file.exists(paste0("RData/", filename.root, "_loo.rds"))){
          jm <- results.all$jm
          
          loo.out[[c]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                        MCMC.params = MCMC.params, 
                                        data.vector = as.vector(jags.data$y))
          saveRDS(loo.out[[c]], file = paste0("RData/", filename.root, "_loo.rds"))
        } else {
          loo.out[[c]] <- readRDS(file = paste0("RData/", filename.root, "_loo.rds"))
          
        }
        
        c <- c + 1
        
      } else {
        print(paste0("Model file ", model.name, " does not exist" ))
      }    
      
    }
    
  }
}

```

Compare the results using LOOIC:

```{r}

looic.esimates <- lapply(lapply(loo.out, 
                                FUN = function(x) x$loo.out),
                         FUN = function(x) x$estimates)

looic <- unlist(lapply(looic.esimates, 
                       FUN = function(x) x["looic", "Estimate"]))

best.model.ID <- which(looic == min(looic))
best.model <- filenames[best.model.ID]
```


According to the LOOIC values, normal process and normal observation models with independent slope and variance terms was considered best. Pareto k values don't look good. 

```{r}
pareto.k <- lapply(loo.out, 
                   FUN = function(x) x$loo.out)
pareto.k.best <- pareto.k[[best.model.ID]]

pareto.k.best
```

Pareto k values are bad. Interesting that the worst model, according to LOOIC, is the best according to Pareto k.

```{r}
pareto.k[[4]]
```

Take a look at all data points vs. pareto k

```{r}

data.jags.W$data.1 %>% select(Frac.Year, Nests, Season) %>%
  mutate(Loc = "W") %>%
  # transmute(Frac.Year = Frac.Year, 
  #           Nests.W = Nests) %>%
  na.omit() -> data.loo

data.loo$pareto.k <- pareto.k.best$diagnostics$pareto_k

ggplot(data = data.loo) + 
  geom_point(aes(x = Frac.Year, y = pareto.k))
```


The models without the Xs showed better Pareto k values (almost all k < 0.5). Compare LOOIC values for those models. 

```{r}
looic[c(2,4,6,8)]
```

They are practically all the same... 


```{r}
v2.best.ID <- 2

jags.out <- readRDS(file = paste0("RData/", filenames[v2.best.ID], ".rds"))

#jags.out <- readRDS(file = paste0("RData/", best.model[[1]], ".rds"))
jm <- jags.out$jm

bayesplot::mcmc_dens(jm$samples, c("theta"))
```

```{r}
data.frame(jm$summary) %>% rownames_to_column("Parameter") -> summary.df

#summary.df[grep(summary.df$Parameter, pattern = "U\\["),]
summary.df[grep(summary.df$Parameter, pattern = "theta"),]
```


```{r}
bayesplot::mcmc_dens(jm$samples, c("sigma.N", "sigma.obs"))
```

Look good.

```{r}
bayesplot::mcmc_trace(jm$samples, c("sigma.N", "sigma.obs"))
```


```{r}
summary.df[grep(summary.df$Parameter, pattern = "sigma."),]

```

