---
title: "Summarizing output from imputation-trend analysis results - Step 2"
output: html_notebook
---

This document summarizes output from imputation/trend analysis results.  The first step should have been done in time_series_trend_Four_summary.Rmd and results saved in RData/time_series_trend_Four_summary_yyyy-mm-dd.rds, where yyyy-mm-dd is the run date. The rds file contains loo output and Pareto k statistics plots for all models. 

```{r}
rm(list=ls())

source('Dc_Indonesia_nesting_fcns.R')

library(jagsUI)
library(coda)
library(tidyverse)
library(loo)

run.date <- "2019-07-01"
```

First load the results:
```{r}
all.results <- readRDS(file = paste0("RData/time_series_trend_Four_summary_", 
                                     run.date,  ".rds"))

```

Get all pareto dataframes together and make a plot:

```{r}
pareto.df <- do.call(rbind, all.results$pareto.df)

p.JM <- ggplot(data = filter(pareto.df, loc == 1)) + 
  geom_path(aes(x = datapoint, y = y), 
            alpha = 0.5) +
  geom_point(aes(x = datapoint, 
                 y = y,
                 color = k0.7)) +
  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))+
  scale_color_manual(values = c("<=0.7" = "black",
                                ">0.7" = "red")) + 
  labs(x = "", y = "", color = "Pareto k") + 
  facet_wrap(model ~., ) + 
  theme(strip.text = element_text(size = 7))

p.JM
```

For JM, either observation model (normal or Student's t), single process slope (U) and process variance (Q), independent Us and single Q models worked well, with respect to Pareto k statistic. 

PLot also for Wermon

```{r}
p.W <- ggplot(data = filter(pareto.df, loc == 2)) + 
  geom_path(aes(x = datapoint, y = y), 
            alpha = 0.5) +
  geom_point(aes(x = datapoint, 
                 y = y,
                 color = k0.7)) +
  # scale_size_continuous(limits = c(0.0, 1.3),
  #                       range = c(1, 4))+
  scale_color_manual(values = c("<=0.7" = "black",
                                ">0.7" = "red")) + 
  labs(x = "", y = "", color = "Pareto k") + 
  facet_wrap(model ~., ) + 
  theme(strip.text = element_text(size = 7))

p.W

```

For Wermon data, either observation model (normal of Student's t) with month-specific process variance models, especially constant CV and discrete Fourier models, and independent slope (U) worked well. 

Compare LOOIC as well?

```{r}
LOOIC <- unlist(lapply(all.results$loo.out, function(x) x$estimates["looic", "Estimate"]))

```

Again, the best LOOIC value goes to one of the worst models based on Pareto k statistic... 


