---
title: "Trend analysis"
output: html_notebook
---

```{r}
rm(list=ls())
library(jagsUI)
library(coda)
library(tidyverse)
source("Dc_Indonesia_nesting_fcns.R")

# the date jags was run (run_JM_SSAR1_logY.R and run_W_SSAR!_logY.R)
run.date <- "2019-03-14"

# MCMC setup - comments are from Summer Martin. Thinning isn't really necessary... but
# good to reduce the output file size. 
n.samples <- 250000   
mcmc.chains <- 5     # 2 is min; AY uses 3
mcmc.thin <- 50      # 50-100 is more than safe; 500 seems excessive per AY
mcmc.burn <- 5000    # 1000; also bumped up to 5000 (same as n.samples) per CB after geweke.diag for deviance in chain 1 still looked high at 2.2
#samples2Save <- (mcmc.burn + n.samples) * mcmc.thin

MCMC.params <- list(n.chains = mcmc.chains,
                    n.samples = n.samples,
                    n.burnin = mcmc.burn,
                    n.thin = mcmc.thin)

# analysis constants
# years to be analyzed for Jamursba-Medi and Wermon
begin.year <- 2001
end.year <- 2018

JM.yrs <- 2001:2018
W.yrs <- 2006:2018

JM.bestM <- 8
W.bestM <- 7
```


This document describes trend analysis of leatherback turtles at nesting beaches in Indonesia. Bayesian state-space time series analyses, which were developed for HI longline fisheries Biological Opinions, were used to estimate the population growth rates.  Code was obtained from Summer Martin, whose code was based on Boyd et al. (2016?)

Data are annual number of females or nests. These data are from the BiOp analysis and will be changed in the future.  The numbers of females are derived from the numbers of nests by dividing them with a fixed clutch frequency so results would not be different between the two.  For reducing a layer, which would require another level of uncertainty to properly analyze these data, I will use the number of nests. 
```{r}
# JM <- read.csv("data/JM_annual_estimated_counts_v2.csv", 
#                header=T)  # Jamursba Medi (1 peak nesting in summer)


JM <- read.csv(paste0("data/estimatedX_JM_M", JM.bestM, ".csv"), 
                header=T)
JMdat <- JM[which(JM$season %in% JM.yrs), ]


W <- read.csv(paste0("data/estimatedX_W_M", W.bestM, ".csv"), 
              header=T)  # Jamursba Medi (1 peak nesting in summer)

Wdat <- W[which(W$season %in% W.yrs), ]   # only keep specific data years 
Wdat[which(Wdat$season %in% 2013:2015) , 
     c("Summer.median", "Winter.median", "all.median")] <- NA  # for seasons 2013-2015, make data = NA since not using; retains time sequence

# 3. COMBINE JM & W into one dataset with 2 time series and run model on that
thedata <- left_join(JMdat, Wdat, by = "season") %>%
  transmute(Season = season, 
            Summer.median.JM = Summer.median.x, 
            Winter.median.JM = Winter.median.x,
            All.median.JM = all.median.x,
            Summer.median.W = Summer.median.y, 
            Winter.median.W = Winter.median.y,
            All.median.W = all.median.y,
            Summer.observed.JM = summer.x, 
            Winter.observed.JM = winter.x,
            All.observed.JM = summer.x + winter.x,
            Summer.observed.W = summer.y, 
            Winter.observed.W = winter.y,
            All.observed.W = summer.y + winter.y)

```

Once data are filtered, set data up for analysis.

First one is Jamusrba-Medi and Wermon combined with a common trend (U) and one process variance (Q)
```{r}
# scenario <- "Dc_singleUQ_Annual_JM_W_Summer_Winter"
# outfile <- paste0("RData/", scenario, ".rds")
# file.tag <- paste0(scenario,"_", Sys.Date(),"_")
#   
# thedata <- thedata.leathers
# modl <- "singleUQ"
# data.type <- "Annual_Females"
# pop.name <- c("Jamursba Medi Leatherback Turtles", "Wermon Leatherback Turtles")
# pop.name.combo <- "Western Pacific Leatherback Turtles"

```

We need to 
```{r}
# Transform data into natural log (Ln) space and PLOT DATA
#log.data <- log(thedata[, c("median_JM", "median_W")])     
log.data <- log(thedata[, c("All.median.JM", "All.median.W")])     
# Ln(TOTAL NUMBER OF FEMALES)
dat <- t(log.data)    # tranpose to have rows = time series and cols = years of data

# Note: The initial state, x0, is treated as a model parameter. 
# the prior matters for Wermon - setting mean to first JM data point skews Wermon to JM trend due to poor W data
# making it specific to each time series for independentUQ


```

```{r}
  
if(!file.exists(paste0("Rdata/singleUQ_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("A", "U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_singleUQ.txt"
  
  # Run model
  seed <- 132
  
  jm.out <- run.singleUQ(dat, 
                         jags.params, 
                         model.loc, 
                         MCMC.params, 
                         seed)
  
  jm.out$data <- thedata
  saveRDS(jm.out, file = paste0("Rdata/singleUQ_out_", 
                                begin.year, "_", end.year,  "_", run.date, ".rds"))
  print(jm.out$out)

} 

```


In the second model, both U (slope) and Q (process variance) are independent between the two time series. 
```{r}
# Model 2:  'independentUQs' ... multiple time series -> multiple population processes (trends)

if(!file.exists(paste0("Rdata/independentUQ_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- paste("models/model_independentUQs.txt", sep="")
  
  # Run model
  set.seed <- 132

  jm.out <- run.independentUQ(dat, 
                              jags.params, 
                              model.loc, 
                              MCMC.params, 
                              seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentUQ_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 

```

Model 3 has single slope (U) and location specific process variance (Q).

```{r}
if(!file.exists(paste0("Rdata/independentQs_singleU_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentQs_singleU.txt"
  
  # Run model
  set.seed <- 132
  jm.out <- run.independentQs_singleU(dat, 
                                      jags.params, 
                                      model.loc, 
                                      MCMC.params, 
                                      seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentQs_singleU_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```

Model 4 has single process variance (Q) and location specific slope (U).

```{r}
if(!file.exists(paste0("Rdata/independentUs_singleQ_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentUs_singleQ.txt"
  
  # Run model
  set.seed <- 132
  jm.out <- run.independentUs_singleQ(dat, 
                                      jags.params, 
                                      model.loc, 
                                      MCMC.params, 
                                      seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentUs_singleQ_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```


Once the four models are run, we compare these using DIC. I haven't figured out how to use looic for this dataset because of missing data points and multi-dimensional likelihood values.  I think I can just treat them as independent values but that requires converting sizes of the log likelihood matrix... so, that's not done yet.

```{r}
tmp <- readRDS(paste0("RData/singleUQ_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U1Q1.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentQs_singleU_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U1Q2.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentUQ_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U2Q2.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentUs_singleQ_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U2Q1.DIC <- tmp$out$DIC

c(U1Q1.DIC, U1Q2.DIC, U2Q2.DIC, U2Q1.DIC)

#minimum is U1Q1
```

Using the simplest model results in the lowest DIC value. So, use that for the following seciton.

```{r}
jags.model <- readRDS(paste0("RData/singleUQ_out_", 
                             begin.year, "_", end.year, "_", run.date, ".rds"))

# create a dataframe for plotting - need median, upper, and lower 95% limits for each 
# pop
plot.df.1 <- data.frame(median = jags.model$out$q50$X,
                        lower = jags.model$out$q2.5$X,
                        upper = jags.model$out$q97.5$X,
                        data = jags.model$jags.data$Y[1,],
                        time = thedata$Season,
                        loc = 1)

p.1 <- ggplot(data = plot.df.1) + 
  geom_ribbon(aes(x = time, ymin = lower, ymax = upper),
              fill = "lightblue") +
  geom_line(aes(x = time, y = median),
            color = "darkblue") + 
  geom_point(aes(x = time, y = data),
             color = "black") +
  labs(x = "", y = "ln(nests)", title = "Jamursba-Medi")

p.1

```

Then for Wermon
```{r}
plot.df.2 <- data.frame(median = jags.model$out$q50$X + jags.model$out$q50$A[2],
                        lower = jags.model$out$q2.5$X + jags.model$out$q2.5$A[2],
                        upper = jags.model$out$q97.5$X + jags.model$out$q97.5$A[2],
                        data = jags.model$jags.data$Y[2,],
                        time = thedata$Season,
                        loc = 2)

p.2 <- ggplot(data = plot.df.2) + 
  geom_ribbon(aes(x = time, ymin = lower, ymax = upper),
              fill = "lightblue") +
  geom_line(aes(x = time, y = median),
            color = "darkblue") + 
  geom_point(aes(x = time, y = data),
             color = "black") +
  labs(x = "", y = "ln(nests)", title = "Wermon")

print(p.2)
```

Look at the posterior of the slope parameter (U).
```{r}
bayesplot::mcmc_dens(jags.model$out$samples, "U")
```

Majority (%) of the posterior is less than 0, indicating a decline. 

Next, we split into summer and winter.
```{r}
# 4. Summer/Winter LEATHERBACKS (again, Tomo's estimates - we are only using medians for now)

# Transform data into natural log (Ln) space and PLOT DATA
log.data <- log(thedata[, c("Summer.median.JM", "Summer.median.W")])   
# Ln(TOTAL NUMBER OF FEMALES)
dat <- t(log.data)    # tranpose to have rows = time series and cols = years of data

```

Then run the three models on just summer and winter separately.

```{r}
if(!file.exists(paste0("Rdata/singleUQ_summer_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("A", "U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_singleUQ.txt"
  
  # Run model
  seed <- 132
  
  jm.out <- run.singleUQ(dat, 
                         jags.params, 
                         model.loc, 
                         MCMC.params, 
                         seed)
  
  jm.out$data <- thedata
  saveRDS(jm.out, file = paste0("Rdata/singleUQ_summer_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)

} 

```

```{r}
if(!file.exists(paste0("Rdata/independentUQ_summer_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentUQs.txt"
                     
  # Run model
  set.seed <- 132

  jm.out <- run.independentUQ(dat, 
                              jags.params, 
                              model.loc, 
                              MCMC.params, 
                              seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentUQ_summer_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```

```{r}
if(!file.exists(paste0("Rdata/independentQs_singleU_summer_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentQs_singleU.txt"
  
  # Run model
  set.seed <- 132
  jm.out <- run.independentQs_singleU(dat, 
                                      jags.params, 
                                      model.loc, 
                                      MCMC.params, 
                                      seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentQs_singleU_summer_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```

```{r}
if(!file.exists(paste0("Rdata/independentUs_singleQ_summer_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentUs_singleQ.txt"
  
  # Run model
  set.seed <- 132
  jm.out <- run.independentUs_singleQ(dat, 
                                      jags.params, 
                                      model.loc, 
                                      MCMC.params, 
                                      seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentUs_singleQ_summer_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```


```{r}
tmp <- readRDS(paste0("RData/singleUQ_summer_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U1Q1.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentQs_singleU_summer_out_", 
                      begin.year, "_", end.year, "_", run.date,  ".rds"))
U1Q2.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentUQ_summer_out_", 
                      begin.year, "_", end.year, "_", run.date,  ".rds"))
U2Q2.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentUs_singleQ_summer_out_", 
                      begin.year, "_", end.year, "_", run.date,  ".rds"))
U2Q1.DIC <- tmp$out$DIC

c(U1Q1.DIC, U1Q2.DIC, U2Q2.DIC, U2Q1.DIC)

#minimum is U1Q1
```

Using the simplest model results in the lowest DIC value. So, use that for the following seciton.

```{r}
jags.model <- readRDS(paste0("RData/singleUQ_summer_out_", 
                             begin.year, "_", end.year, "_", run.date, ".rds"))

# create a dataframe for plotting - need median, upper, and lower 95% limits for each 
# pop
plot.df.1 <- data.frame(median = jags.model$out$q50$X,
                        lower = jags.model$out$q2.5$X,
                        upper = jags.model$out$q97.5$X,
                        data = jags.model$jags.data$Y[1,],
                        time = thedata$Season,
                        loc = 1)

p.1 <- ggplot(data = plot.df.1) + 
  geom_ribbon(aes(x = time, ymin = lower, ymax = upper),
              fill = "lightblue") +
  geom_line(aes(x = time, y = median),
            color = "darkblue") + 
  geom_point(aes(x = time, y = data),
             color = "black") +
  labs(x = "", y = "ln(nests)", title = "Jamursba-Medi (Summer)")

p.1

```

```{r}
plot.df.2 <- data.frame(median = jags.model$out$q50$X + jags.model$out$q50$A[2],
                        lower = jags.model$out$q2.5$X + jags.model$out$q2.5$A[2],
                        upper = jags.model$out$q97.5$X + jags.model$out$q97.5$A[2],
                        data = jags.model$jags.data$Y[2,],
                        time = thedata$Season,
                        loc = 2)

p.2 <- ggplot(data = plot.df.2) + 
  geom_ribbon(aes(x = time, ymin = lower, ymax = upper),
              fill = "lightblue") +
  geom_line(aes(x = time, y = median),
            color = "darkblue") + 
  geom_point(aes(x = time, y = data),
             color = "black") +
  labs(x = "", y = "ln(nests)", title = "Wermon (Summer)")

print(p.2)
```

Look at the posterior of the slope parameter (U).
```{r}
bayesplot::mcmc_dens(jags.model$out$samples, "U")
```

80% of the posterior is less than 0. 

Then winter

```{r}
# 4. Summer/Winter LEATHERBACKS (again, Tomo's estimates - we are only using medians for now)

# import both summer and winter data files (each has separate time series within for JM and W)
# Transform data into natural log (Ln) space and PLOT DATA
#log.data <- log(thedata[, c("median_JM", "median_W")])     
log.data <- log(thedata[, c("Winter.median.JM", "Winter.median.W")])   
#
# Ln(TOTAL NUMBER OF FEMALES)
dat <- t(log.data)    # tranpose to have rows = time series and cols = years of data

```

Then run the three models on just summer and winter separately.

```{r}
if(!file.exists(paste0("Rdata/singleUQ_winter_out_", 
                       begin.year, "_", end.year, "_", run.date,  ".rds"))){
  jags.params <- c("A", "U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_singleUQ.txt"
  
  # Run model
  seed <- 132
  
  jm.out <- run.singleUQ(dat, 
                         jags.params, 
                         model.loc, 
                         MCMC.params, 
                         seed)
  
  jm.out$data <- thedata
  saveRDS(jm.out, file = paste0("Rdata/singleUQ_winter_out_", 
                                begin.year, "_", end.year,  "_", run.date, ".rds"))
  print(jm.out$out)

} 

```

```{r}
if(!file.exists(paste0("Rdata/independentUQ_winter_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentUQs.txt"
                     
  # Run model
  set.seed <- 132

  jm.out <- run.independentUQ(dat, 
                              jags.params, 
                              model.loc, 
                              MCMC.params, 
                              seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentUQ_winter_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```

```{r}
if(!file.exists(paste0("Rdata/independentQs_singleU_winter_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentQs_singleU.txt"
  
  # Run model
  set.seed <- 132
  jm.out <- run.independentQs_singleU(dat, 
                                      jags.params, 
                                      model.loc, 
                                      MCMC.params, 
                                      seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentQs_singleU_winter_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```

```{r}
if(!file.exists(paste0("Rdata/independentUs_singleQ_winter_out_", 
                       begin.year, "_", end.year, "_", run.date, ".rds"))){
  jags.params <- c("U", "Q", "R", "X0", "X", "loglik")
  model.loc <- "models/model_independentUs_singleQ.txt"
  
  # Run model
  set.seed <- 132
  jm.out <- run.independentUs_singleQ(dat, 
                                      jags.params, 
                                      model.loc, 
                                      MCMC.params, 
                                      seed)
  
  jm.out$data <- thedata
  
  saveRDS(jm.out, file = paste0("Rdata/independentUs_singleQ_winter_out_", 
                                begin.year, "_", end.year, "_", run.date, ".rds"))
  print(jm.out$out)
} 
```



```{r}
tmp <- readRDS(paste0("RData/singleUQ_winter_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U1Q1.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentQs_singleU_winter_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U1Q2.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentUQ_winter_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U2Q2.DIC <- tmp$out$DIC
tmp <- readRDS(paste0("RData/independentUs_singleQ_winter_out_", 
                      begin.year, "_", end.year, "_", run.date, ".rds"))
U2Q1.DIC <- tmp$out$DIC

c(U1Q1.DIC, U1Q2.DIC, U2Q2.DIC, U2Q1.DIC)

#minimum is U1Q1
```


```{r}
jags.model <- readRDS(paste0("RData/singleUQ_winter_out_", 
                             begin.year, "_", end.year, "_", run.date, ".rds"))

# create a dataframe for plotting - need median, upper, and lower 95% limits for each 
# pop
plot.df.1 <- data.frame(median = jags.model$out$q50$X,
                        lower = jags.model$out$q2.5$X,
                        upper = jags.model$out$q97.5$X,
                        data = jags.model$jags.data$Y[1,],
                        time = thedata$Season,
                        loc = 1)

p.1 <- ggplot(data = plot.df.1) + 
  geom_ribbon(aes(x = time, ymin = lower, ymax = upper),
              fill = "lightblue") +
  geom_line(aes(x = time, y = median),
            color = "darkblue") + 
  geom_point(aes(x = time, y = data),
             color = "black") +
  labs(x = "", y = "ln(nests)", title = "Jamursba-Medi (Winter)")

p.1

```

```{r}
plot.df.2 <- data.frame(median = jags.model$out$q50$X + jags.model$out$q50$A[2],
                        lower = jags.model$out$q2.5$X + jags.model$out$q2.5$A[2],
                        upper = jags.model$out$q97.5$X + jags.model$out$q97.5$A[2],
                        data = jags.model$jags.data$Y[2,],
                        time = thedata$Season,
                        loc = 2)

p.2 <- ggplot(data = plot.df.2) + 
  geom_ribbon(aes(x = time, ymin = lower, ymax = upper),
              fill = "lightblue") +
  geom_line(aes(x = time, y = median),
            color = "darkblue") + 
  geom_point(aes(x = time, y = data),
             color = "black") +
  labs(x = "", y = "ln(nests)", title = "Wermon (Winter)")

print(p.2)
```

Look at the posterior of the slope parameter (U).
```{r}
bayesplot::mcmc_dens(jags.model$out$samples, "U")
```


X% of U is < 0. 
